\documentclass{article}
% \usepackage{showframe}

\usepackage{../../preamble-notes}

% Custom Definitions of operators
\DeclareMathOperator{\Ima}{im}
\DeclareMathOperator{\Fix}{Fix}
\DeclareMathOperator{\Orb}{Orb}
\DeclareMathOperator{\Stab}{Stab}
\DeclareMathOperator{\send}{send}
\DeclareMathOperator{\dom}{dom}
\DeclareMathOperator{\Int}{int}
\DeclareMathOperator{\id}{id}

\title{General Topology Math Notes}
\author{Leon Lee}
\renewcommand\labelitemi{\tiny$\bullet$}

\begin{document}

\maketitle
\newpage
\tableofcontents
\newpage

\section{Intro to Topology}
\subsection{Why Topology?}

Topology can appear where we least expect it...
\begin{itemize}
    \item Algebraic Number Theory - Next to Euclidean topology, can define other topologies on $\mathbb{Q}$ (related to how often primes divide a number). Extends to Adeles, Langlands programme, etc
    \item Arithmetic Progressions in the Integers - An arithmetic progression of length $k$ is a set $\{a, a+d,\dots,a+(k-1)d\}$
        Finding subsets of $\mathbb{N}$ that contain arbitrarily long APs:
        \begin{itemize}
            \item $2\mathbb{N}$ or $\mathbb{N}$
            \item Primes (Green-Tao Theorem, 2007). Green-Tao theorem relies on \textbf{Szemeredi's Theorem}: Any dense enough subset of $\mathbb{N}$ contains arbitrarily long APs

                Furstenburg's idea: Get from $A \subseteq\mathbb{N}$ to $(a_{i}\in \{0, 1\}^{\mathbb{N}})$ with $a_{i} \begin{cases}
                    1 & i\in A \\
                    0 & \text{else}
                \end{cases}$

                Use topological dynamics to study this: A topological dynamical system is a triple of $X$ cpt, $T : X \to X$ continuous, and a probability measure $\mu$ preserved by $T$ (what)
        \end{itemize}
\end{itemize}

\subsection{Topological Spaces and Examples}
\begin{dfn}[Topological Space]{dfn:top-space}{}
    A \textbf{topological space} is a pair $(X, \mathcal{T})$, where $X$ is a nonempty set, and $\mathcal{T}$ is a collection of subsets of $X$ which satisfies:
    \begin{enumerate}
        \item $\emptyset\in \mathcal{T}$ and $X\in \mathcal{T}$
        \item if $U_{\lambda}\in \mathcal{T}$ for each $\lambda\in A$ (where $A$ is some indexing set), then $\bigcup\limits_{\lambda \in A} U_{\lambda}\in \mathcal{T}$
        \item If $U_{1}, U_{2}\in \mathcal{T}$, then $U_{1} \cap U_{2}\in \mathcal{T}$
    \end{enumerate}
\end{dfn}

\subsubsection{Examples of Topological Spaces}
\begin{enumerate}
    \item $\mathbb{R}^{n}$ with the Euclidean Topology - induced by the Euclidean Metric
    \item For any set $X$, $\mathcal{T} = \mathcal{P}(X)$ (discrete topology)
    \item For any set $X$, $\mathcal{T} = \{\emptyset, X\}$ (indiscrete topology)
    \item $X = \{0,1,2\}$ with $\mathcal{T} = \{\emptyset, X, \{0\}, \{0,1\}, \{0,2\}\}$
    \item $X = \mathbb{R}$ and $U$ open (aka, in $\mathcal{T}$) if $R \backslash U$ is finite or $U = \emptyset$
\end{enumerate}
Proof for $5$:
\begin{enumerate}
    \item $\emptyset\in \mathcal{T}$, $\emptyset$ is finite $\implies X \in \mathcal{T}$
    \item Intersections of finite sets are finite
    \item Unions of finite sets are finite
\end{enumerate}

\begin{dfn}[Neighbourhood of a point]{dfn:neighbourhood}{}
    A \textbf{neighbourhood} of a point $x\in X$ is a subset $N \subseteq X$ s.t. $x\in U \subseteq N$ for some open subset $U \subseteq X$
\end{dfn}

\begin{dfn}[Metric Space]{dfn:metric-space}{}
    A \textbf{metric space} $(X, d)$ is a nonempty set $X$ together with a function
    \[d : X \times X \to \mathbb{R}\]
    with the following properties:
    \begin{enumerate}
        \item $d(x, y) \ge 0$ for all $x, y\in X$ and $d(x, y) = 0$ iff $x = y$
        \item $d(x, y) = d(y, x)$ for all $x, y \in X$
        \item $d(x, z) \le d(x, y) + d(y, z)$ for all $x,y,z\in X$
    \end{enumerate}
    The function $d$ is called the metric. Point $3$ is called the \textit{triangle inequality}

    \longrule{0.08ex}
    
    For any $x\in X$ and any positive real number $r$ the \textbf{open ball} in $X$ with centre $x$ and radius $r$ is defined by
    \[B(x, r) = \{y\in X | d(x, y) < r\}\]
    We declare a subset $U$ of $X$ to be \textit{open in the metric topology given by d} iff for each $a\in U$ there is an $r > 0$ such that $B(a, r) \subseteq U$

    \longrule{0.08ex}

    If $(X, \mathcal{T})$ is a topological space, and if $X$ admits a metric whose metric topology is precisely $\mathcal{T}$ we say that $(X, \mathcal{T})$ is \textbf{metrisable}. Thus the euclidean spaces with their usual topologies are metrisable
\end{dfn}

\subsubsection{Examples of Metric Spaces}
\begin{enumerate}
    \item Any set $X$ with $d(x, y) = \begin{cases}
            0, & x = y \\
            1, & \text{else}
    \end{cases}$
    \item $\mathbb{R}^{n}$ with $d(x, y) = \lvert x - y \rvert = \sqrt{\sum_{i = 1}^{n} (x_{i} - y_{i})^{2}}$
    \item $C([0,1])$ with $d(f, g) = \max_{t\in[0,1]} \lvert f(t) - g(t) \rvert$
    \item $C([0,1])$ with $d(f, g) = \sqrt{\int_{0}^{n}\lvert f(t) - g(t) \rvert^{2} dt}$
\end{enumerate}

\subsubsection{Topologies on Metric spaces}
We want to define a topology on $(X, d)$. For this, we want open balls to be open in the topology

\begin{dfn}[Base]{dfn:base}{}
    For a set $X$, a basis $\mathcal{B}$ is a collection of subsets such that
    \begin{enumerate}
        \item $\bigcup_{B\in\mathcal{B}} B = X$
        \item $B_{1}\cap B_{2}\in \mathcal{B}$ for all $B_{1},B_{2}\in \mathcal{B}$
    \end{enumerate}
    The \textbf{topology generated by $\mathcal{B}$} is
    \[\mathcal{T} := \{\bigcup_{i\in I} B_{i}, \text{$I$ index set},\, B_{i}\in \mathcal{B}\}\]
\end{dfn}

\textbf{Note}: This is a topology because
\[(\cup_{i\in I} B_{i}) \cap (\cup_{j\in J} B_{j}) = \bigcup_{i\in I,j\in J} \underbrace{B_{i} \cap B_{j}}_{\in \mathcal{B}}\in \mathcal{t}\]

\begin{dfn}[Metric Topology]{dfn:metric-topology}{}
    \[\text{Let } \mathcal{B} = \{\bigcap_{i = 1}^{n} Br_{1}(x_{1}), n\in \mathbb{N},\, r_{i} > 0,\, x_{i}\in X,\,\forall i\}\]
    The \textbf{metric topology} is the topology generated by this basis
\end{dfn}

\textbf{Observation} A set $U$ is open in the metric topology $\iff$ $\forall x\in U,\, \exists r > 0$ s.t. $Br(x) \subseteq U$

\begin{itemize}
    \item $\impliedby$: For each $x\in U$, let $r_{x}$ s.t. $B_{r_{x}}(x) \subseteq U$. Then $U = \bigcup\limits_{x \subseteq U} B_{r_{x}}(x)$ is open
    \item $\implies$: Let $x\in U$ be given. Knwo that $x\in B_{r_{1}}(x_{1}) \cup \dots \cup B_{r_{n}}(x_{n})$ for some $n,r_{1},x_{1}$. For each $i$, there is $\delta_{i} >0$ s.t. $B_{\delta_{i}}(x) \le B_{r_{1}}(x_{1})$.
\end{itemize}

huh?

\begin{thm}[random ms prop]{thm:ms-prop-idk}{}
    If $X$ carries metrics $d, \tilde{d}$ such that $ad(x, y) \le \tilde{d}(x,y) \le Ad(x, y)$ for some $a, A > 0$, then the induced topologies agree
\end{thm}


\begin{dfn}[Subspace Topology]{dfn:subspace-topology}{}
    Let $(X, \mathcal{T})$ be a topological space, and let $A \subseteq X$ be any subset. Then the \textbf{subspace topology} on $A$ consists of all sets of the form $U \cap A$ where $U \in \mathcal{T}$
\end{dfn}

\textbf{Example}: $(-1, 1) \subseteq \mathbb{R}$ with euclidean topology. The subspace topology is
\[\{(-1, 1) \cap U,\, U \subseteq \mathbb{R} \text{ open}\}\]
$(-1, 1)$ is closed in the subspace topology

\begin{thm}[Topology Lemmas]{thm:topology-lemmas}{}
    \begin{itemize}
        \item[\textbf{1.3}] If $(X, \mathcal{T})$ is a topological space and $U_{1},\dots,U_{n}$ are open sets, then the intersection $\bigcap_{i=1}^{n}u_{i}$ is also open
        \item[1.6] In order to show that a set $U \subseteq X$ is open, it is enough to show that for every $x\in U$ there is an open set $V$ with $x\in V \subseteq U$
        \item[1.6] A subset $U$ of $\mathbb{R}^{n}$ is \textit{open for the usual topology} iff for each $a\in U$ there exists an $r > 0$ s.t.
            \[\lvert x -a \rvert < r \implies x \in U\]
            The collection of open sets thus defined is called the \textbf{usual topology on $\mathbb{R}^{n}$}. Note that open balls are open sets under this definition
    \end{itemize}
\end{thm}

\begin{dfn}[Topology Small Definitions]{dfn:topology-small-defs}{}
    \begin{itemize}
        \item
    \end{itemize}
\end{dfn}


\newpage

\subsection{Closed sets, Closure, Interior, and Boundary}

\begin{dfn}[Closed Subsets]{dfn:closed-subset}{}
    Let $(X, \mathcal{T})$ be a topological space. A subset $A \subseteq X$ is \textbf{closed} iff its complement $X \backslash A = A^{C} := \{x \in X \ x\not\in A\}$ is open in $X$
\end{dfn}

\textbf{Note}: A set being ``closed'' has no connection with ``not being open''

\subsubsection{Examples of open and closed sets}
\begin{itemize}
    \item A set that is neither open nor closed: $[0, 1) \subseteq \mathbb{R}$ under Euclidean topology
    \item A set that is both closed and open: $\emptyset$ or $X$
\end{itemize}

\begin{thm}[]{thm:topology-union-lemmas}{}
    Let $(X, \mathcal{T})$ be a topological space. Then
    \begin{enumerate}
        \item $\emptyset$ and $X$ are closed.
        \item The union of finitely many closed sets is a closed set
        \item The intersection of any collection of closed sets is a closed set
    \end{enumerate}
\end{thm}

$\bigcup_{i\in I} A_{i}$ is not necessarily closed.
\[A_{n} = \left[-2 + \frac{1}{n}, 2 - \frac{1}{n}\right] \text{ in } \mathbb{R}, \bigcup_{i = 1}^{n} A_{i} = (-2, 2)\]

\subsubsection{Topological proof that there are infinitely many primes (Furstenberg)}
\begin{proof}
    Look at $\mathbb{Z}$ with

    \[\mathcal{B} := \{S(a, b),\, a\ne 0,\,b\in \mathbb{Z}\} \quad\text{ and }\quad S(a,b) = \{an + b,\,n\in \mathbb{Z}\}\]

    Let the open sets be the one generated by this basis. We can show
    \begin{enumerate}
        \item $S(a,b)$ is both open and closed.
        \item All open sets are infinite.
    \end{enumerate}

    \vspace{-3pt}
    \noindent\rule{\textwidth}{0.08ex}\vspace{-8pt}

    \begin{enumerate}
        \item $S(a,b) = \mathbb{Z} \backslash \bigcup_{i=1}^{a - 1} S(a, b-1)$
        \item Clear
    \end{enumerate}

    Thus:
    \[\underbrace{\mathbb{Z} \backslash \{-1, 1\}}_{\text{not closed}} = \bigcup_{\text{$p$ primes}} \overbrace{S(p, 0)}^{\text{closed}}\]
    If there were only finitely many primes, right hand side would be closed
\end{proof}

\newpage
\subsection{Closure and stuff}

\begin{dfn}[Closure, Interior, Boundary]{dfn:closure-interior-boundary}{}
    Let $(X, \mathcal{T})$ be a topological space.
    \begin{enumerate}
        \item The \textbf{closure} of a subset $A \subseteq X$ is the smallest closed set such that $A \subseteq \overline{A}$.
            \[\overline{A} := \bigcap\limits_{\substack{C \subseteq X \text{closed} \\ A \subseteq C}} C\]
        \item The (open) \textbf{interior} of a subset $A \subseteq X$ is the biggest open set $U$ contained in $A$
            \[\Int{A} = A^{\circ} := \bigcap\limits_{\substack{U \subseteq X \text{open} \\ U \subseteq A}} C\]
        \item The \textbf{boundary} or \textbf{frontier} of a subset $A \subseteq X$ is
            \[\partial A := \overline{A} \backslash A^{\circ}\]
        \item A subset $A$ of $X$ is \textbf{dense} in $X$ iff $\overline{A} = X$

            E.g.: $\mathbb{Q} \subseteq \mathbb{R}$ with the Euclidean topology
    \end{enumerate}
\end{dfn}

\begin{thm}[Closure and Interior of Complement]{thm:complement-properties}{}
    Let $(X, \mathcal{T})$ be a topological space and $A \subseteq X$. Then
    \begin{enumerate}
        \item The closure of the complement is the complement of the interior:
            \[\overline{X \backslash A} = X \backslash (A^{\circ})\]
        \item the interior of the complement is the complement of the closure:
            \[(X \backslash A)^{\circ} = X \backslash \overline{A}\]
    \end{enumerate}
\end{thm}

\begin{dfn}[Limits in Topological spaces]{dfn:topological-limits}{}
    A sequence $(x_{n})$ converges to $x\in X$ if $\forall U$ open with $x\in U$, $\exists N$ s.t. $x_{n}\in U$ for all $n \ge N$
\end{dfn}

\begin{dfn}[Limit Set]{dfn:limitset}{}
    $\overline{A}$ can be thought as the set of limits. Formally define the \textbf{Limit Set} as
    \[A' = \{x\in X : \forall U \subseteq X, x\in U, \text{ open } \exists a\in A,\,\text{s.t. } a\in U\}\]
\end{dfn}

But, limits in general are much worse behaved in topological spaces, e.g. no unique limit point

\textbf{Example}: a topological space $X$ and a sequence $(x_{n})$ which does not have a unique limit (i.e. $\exists x\ne y$ s.t. $x_{n} \to x$ and $x_{n}\to y$ in the sense defined): Nontrivial $X$ with the indiscrete topology $\{\emptyset, X\}$


\newpage
\subsection{Hausdorff Spaces}

\textbf{Problem}: Non-unique limits are nasty :(

\begin{dfn}[Hausdorff Space]{dfn:hausdorff}{}
    A topological space $(X, \mathcal{T})$ is \textbf{Hausdorff} if for each $x, y\in X$ with $x\ne y$ there exist \textit{disjoint} open sets $U$ and $V$ s.t. $x\in U$ and $y\in V$

    This space has \textit{unique limits}!
\end{dfn}

If $(X, d)$ is a metric space then it is automatically Hausdorff, so any metrisable space is Hausdorff. The trivial topology on a set with more than one element is not Hausdorff. Not every Hausdorff space is metrisable

Non-Hausdorff spaces are a lot more annoying to work with - for example you can have multiple limits in non-Hausdorff spaces

\begin{thm}[Open sets on \texorpdfstring{$\mathbb{R}$}{R} with Euclidean Topology]{thm:r-open-sets}{}
    \begin{itemize}
        \item A set $U$ is open iff there are open intervals $I_{j}$ s.t.
    \[U = \bigcup_{j=1}^{\infty} I_{j}\]
    \item A set $A$ is closed iff there are $F_{j}$ (union of two closed intervals) s.t.
        \[A = \bigcap_{j = 1}^{\infty} F_{j}\]
    \end{itemize}
\end{thm}


\begin{dfn}[Convergence of Hausdorff Spaces]{dfn:haursdoff-convergence}{}
    A sequence $(x_{n})$ of members of a topological space $X$ converges to $x\in X$ if for every open set $U$ containing $x$, there exists an $N$ such that $n\ge N \implies x_{n}\in U$
\end{dfn}

\begin{thm}[Haussdorf Convergence Uniqueness]{thm:hausdorff-conv-uniqueness}{}
    Suppose $(X, \mathcal{T})$ is Hausdorff. Then a sequence $(x_{n})$ can converge to at most one limit.
\end{thm}

Being Hausdorff is what's called a \textit{topological property}, which means whether it is true in a particular case depends only on the open sets of the space in question.

In contrast, the property of \textit{completeness} of a metric space is not a topological property as there exist sets upon which one can put two distinct metrics, one complete and one not, yet for which the metric topologies coincide

\newpage
\begin{dfn}[Cauchy Sequences]{dfn:cauchy-seq}{}
    Let $(X, d)$ be a metric space
    \begin{enumerate}
        \item A \textbf{Cauchy Sequence} is a sequence $(x_{n})$ with each $x_{n}\in X$ with the property that for each $\epsilon > 0$, there exists an $N$ s.t. $m,n\ge N \implies d(x_{m}, x_{n}) < \epsilon$
        \item $(X, d)$ is \textbf{complete} if every Cauchy Sequence converges
    \end{enumerate}
\end{dfn}

\textbf{Caveat}: In general, this does not have to converge to an $x\in X$

\textbf{Example}: $\mathbb{Q}$ with the Euclidean metric. 

\begin{dfn}[Complete Space]{dfn:complete-space}{}
    A metric space is called \textbf{complete} if all Cauchy sequences converge to a point in the space
\end{dfn}

\begin{dfn}[Closure in Metric Spaces]{dfn:metric-closure}{}
    Let $(X, d)$ be a complete metric space and $A \subseteq X$. A point $x$ is in the \textbf{closure} of $A \iff \exists x_{i} \to x$ with $x\in A$
\end{dfn}

\section{Continuity}

\subsection{Continuity}

\begin{dfn}[Continuity]{dfn:continuity}{}
    Let $(X, \mathcal{T}), (Y, \tilde{\mathcal{T}})$ be topological spaces and $f : X \to Y$. $f$ is \textbf{continuous} if for all $U\in \tilde{\mathcal{T}}$, $f^{-1}(U)\in \mathcal{T}$

    \longrule{0.08ex}

    Equivalently:
    \begin{itemize}
        \item $U \subseteq Y$ open $\implies$ $f^{-1}(U)$ open
        \item $A \subseteq Y$ closed $\implies$ $f^{-1}(A)$ closed
    \end{itemize}
\end{dfn}

\subsubsection{Why take \texorpdfstring{$f^{-1}$}{f-1}}

\textbf{Properties}: For $U, V$ sets in $Y$,
\begin{itemize}
    \item $f^{-1}(U \cap V) = f^{-1}(U) \cap f^{-1}(U)$
    \item $f^{-1}(U^{C}) = f^{-1}(U)^{C}$
    \item $f^{-1}(U \cup U) = f^{-1}(U) \cup f^{-1}(U)$
\end{itemize}

\textbf{Example}: $\mathbb{R}$ with Euclidean Topology
\begin{proof}
    "Proof" that $[-1, 1]$ is open:

    Take $[-1, 1]$ with the subspace topology $\mathcal{T} := \{[-1, 1] \cap U,\, U \subseteq \mathbb{R} \text{ open}\}$

    Embedding $i : [-1, 1] \to \mathbb{R},\,x \mapsto x$ is continuous

    $[-1, 1]$ open in subspace topology

    $i$ cont $\implies$ $i([-1,1])$ is open $\quad$ \textcolor{red}{this is actually wrong! $U$ open $\centernot\implies f(U)$ open}

    But $i([-1, 1]) = [-1, 1] \subseteq \mathbb{R}$
\end{proof}

\begin{dfn}[Formal Definition of Continuity]{dfn:e-d-continuity}{}
    Let $(X, d), (Y, d)$ be metric spaces with the metric topology. $f: X \to Y$ is continuous as above iff $\forall x\in X$, $\forall \epsilon > 0$, $\exists \delta > 0$ s.t.
    \[d(x, y) < \delta \implies d(f(x), f(y)) < \epsilon\]
\end{dfn}

\begin{proof} \textbf{$\implies$ Direction}

    Recall: $U$ open in metric topology if $\forall x\in U,\, \exists r > 0$ s.t. $B_{r}(x) \subseteq U$, where $B_{r}(x) = \{y \in X : d(x, y) < r\}$

    $\implies$ Let $x\in X$ be given, $\epsilon > 0$. Let $y = f(x)\in Y,\, U = B_{\epsilon}(y) = \{y'\in Y : \tilde{d}(y, y') < \epsilon\}$.

    $f$ cont $\implies$ $f^{-1}(U)$ is open. $x\in f^{-1}(U) \implies \exists \delta > 0$ s.t. $B_{\delta}(x) \subseteq f^{-1}(U)$

    $\implies \forall x'\in X$ s.t. $d(x, x') < \delta$, $x'\in B_{\delta}(x) \subseteq f^{-1}(U)$.

    $\implies f(x') \in B_{\delta}(f(x)) \implies \tilde{d}(f(x), f(x')) < \epsilon$

    \textbf{$\impliedby$ Direction}

    Let $U$ be open in $Y$. WTS: $f^{-1}(U)$ is open.

    So it is enough to show for all $x\in f^{-1}(U)$, $\exists \delta > 0$ s.t. $B_{\delta}(x) \subseteq f^{-1}(U)$.

    Let $x$ be given, $y := f(x)\in U$. $U$ open $\implies \exists \epsilon > 0$ s.t. $B_{\epsilon}(y) \subseteq U$.

    By assumption $\exists \delta > 0$ s.t.
    \[d(x', x) < \delta \implies \tilde{d}(f(x'), f(x)) < \epsilon\]
    But, $\{y' : d(y', f(x)) < \epsilon\} \subseteq U$ by choice of $\epsilon$.

    $\implies B_{\delta}(x) \subseteq f^{-1}(U)$
\end{proof}

\subsection{Homeomorphisms}

\begin{dfn}[Homeomorphism]{dfn:homoemorphism}{}
    Let $(X, \mathcal{T}), (Y, \tilde{\mathcal{T}})$ be topological spaces. A function $f : X \to Y$ is a \textbf{homeomorphism} (or \textbf{bi-continuous}) if $f$ is bijective, $f$ is continuous, and $f^{-1} Y \to X$ is continuous
\end{dfn}

A ``Great goal of Topology'': Understand topological spaces up to homeomorphisms.

Say that a property of a topological space is a \textbf{topological invariant} if it is preserved by homeomorphism. Example: Being Hausdorff

\begin{xmp}[Examples of Homeomorphisms]{xmp:homoeomorphism}{}
    \begin{itemize}
        \item $(X, \mathcal{T})$ topological space, $\id : X \to X,\,x \mapsto x$
        \item $X = \mathbb{R}^{n},\, A : \mathbb{R}^{n} \to \mathbb{R}^{n}$ Linear + Invertible
        \item Example which is \textbf{not} a homeomorphism:
            \[f : \underbrace{\mathbb{R}}_{\text{metric topology}} \to \underbrace{\mathbb{R}}_{\text{indiscrete topology $\{\emptyset, \mathbb{R}\}$}},\,x \mapsto x\]

            Problem: $f^{-1}$ is not continuous
    \end{itemize}
\end{xmp}

\newpage

\begin{dfn}[Another continuity definition]{dfn:continuity-2}{}
    Let $(X, d), (Y, \tilde{d})$ be metric spaces with the metric topology. $f : X \to Y$ is continuous iff
    \[\forall x\in X,\,\forall \epsilon > 0,\,\exists \delta > 0\,\, \text{ s.t. } \forall y\in X,\,d(x, y) < \delta \implies \tilde{d}(f(x),f(y)) < \epsilon\]
\end{dfn}

\textbf{Observe}: $\forall y\in X$ is equivalent to
\[B_{\delta}(x) \subseteq f^{-1}(\tilde{B_{\epsilon}}(f(x)))\]

Why? Let $A, B$ be things which can be true for $y\in X$. i.e.
\[A \implies B \text{ is equivalent to } \{y : A \text{ true} \subseteq \{y : B \text{ true}\}\}\]

Then: $B_{\delta}(x) = \{y, \underbrace{d(x,y) < \delta}_{A}\}$, $f^{-1}(\tilde{B_{\epsilon}}(f(x))) = \{y \in X : \underbrace{\tilde{d}(f(x), f(y)) < \epsilon}_{B}\}$

WTS: $U$ open $\iff$ $\forall x\in U,\,\exists r > 0$ s.t. $B_{\delta}(x) \subseteq U$

$\implies$ ``Let $x, \epsilon$ be given, WTS that $\exists \delta$ s.t. $B_{\delta}(x) \subseteq f^{-1}(\tilde{B_{\epsilon}}f(x))$

\textbf{Example}: $f$ cont + bijective but not a homeomorphism:
\[f : \underbrace{X}_{\mathclap{\text{discrete topology - every set is open}}} \to \overbrace{X}^{\mathclap{\text{indiscrete topology: only $\emptyset$ and $X$ are open}}} \text{ identity}\]

\begin{lma}[Homeomorphism-condition]{thm:homeomorphism-condition}{}
    For a set $X$ with topologies $\mathcal{T},\,\tilde{\mathcal{T}}$. The identity map $(X, \mathcal{T})\to (X, \tilde{\mathcal{T}}),\,x \mapsto x$ is
    \begin{itemize}
        \item continuous $\iff$ $\tilde{\mathcal{T}} \subseteq T$
        \item a homeo $\iff$ $\tilde{\mathcal{T}} = \mathcal{T}$
    \end{itemize}
\end{lma}

\begin{thm}[Mapping prop]{thm:mapping-prop}{}
    \begin{itemize}
        \item Let $f : X \to Y,\,g : Y \to Z$ continuous. The map $f \circ g$ is continuous
            \[\text{As } (f \circ g)^{-1}(U) = f^{-1}(g^{-1}(U))\]
        \item If $f : X \to Y$ is constant, then $f$ is continuous
        \item In particular, $f, g$ homeo $\implies$ $f \circ g$ is a homeo
    \end{itemize}
\end{thm}

\newpage

\section{The Clark Barwick Era}

\begin{thm}[Clark Barwick Quotes List]{thm:clark-quotes}{}
    ``Shadows are harshest when there is only one lamp'' - 04/10/24

    ``If one must choose between rigour and meaning, then I shall unhesitantly choose the latter'' - 25/10/24

    ``Everyone knows what a curve is, until they have learned enough mathematics to have become confused'' - F. Klein (29/10/24)

    ``You become what you give your attention to'' - (05/11/24)
\end{thm}

\subsection{More top}
\subsubsection{Something weird}

\begin{align*}
    [0, 2\pi) & \to S^{1} = \{z \in \mathbb{Z} : \lVert z \rVert = 1\}\\
    [0, 2\pi) & \to [0, 1) \quad \text{is open, and is also creepy}
\end{align*}

Not a homeomorphism

\textbf{Claim}: A continuous bijection in which the \textbf{image} of every open set is open is a homeomorphism

\begin{dfn}[Subspace Topology]{dfn:subspace}{}
    For $X$ a topological space, and $T \subseteq X$, $\mathcal{U} \subseteq T$ is open iff $\exists V \subseteq X$ open and $\mathcal{U} = V \cap T$
\end{dfn}

\begin{dfn}[Impromptu Set Theory - Products]{dfn:infinite-products}{}
    $\mathcal{F}$ is a family of sets. We can talk about a product
    \[\bigsqcap_{x\in \mathcal{F}} X = \{(a_{x})_{x\in \mathcal{F}} : a_{X}\in X\}\]
\end{dfn}

Example:
\[\mathbb{R}^{\infty} = \bigsqcap_{i=1}^{\infty} \mathbb{R} = \{(x_{1},x_{2},\dots) : x_{i}\in \mathbb{R}\}\]
\[\bigsqcap_{x\in \mathcal{F}} = \{ \phi : \mathcal{F} \implies \bigcup_{x\in \mathcal{F}} X : \phi(X) \in X\}\]

Note: the $\mathcal{F}$ notation is pretty creepy - Clark

\subsubsection{Topologising the above thing}

\[\bigsqcap_{i\in I} X_{i} \to X_{j}\]

\subsection{Week 4 Lecture 1}

\begin{dfn}[Quotient Topology]{dfn:quotient-top}{}
    Define $X$ with $\sim$ a relation on $X$. We have a function
    \begin{align*}
        g : X &\to X /\sim\\
        x &\mapsto [x]
    \end{align*}
    $\mathcal{U} \subseteq X /\sim$ is open iff $g^{-1}(\mathcal{U}) \subseteq X$ is open
\end{dfn}

\subsection{Week 4 Lecture 2}

\begin{dfn}[Coarser and Finer]{dfn:coarse-fine}{}
    \begin{itemize}
        \item Coarse: There are more open sets
    \end{itemize}
\end{dfn}

\begin{dfn}[Product Topology]{dfn:product-top}{}
    % https://q.uiver.app/#q=WzAsMyxbMSwwLCJYIFxcdGltZXMgWSJdLFswLDEsIlgiXSxbMiwxLCJZIl0sWzAsMSwicHJfMSIsMl0sWzAsMiwicHJfMiJdXQ==
    \[\begin{tikzcd}
    & {X \times Y} \\
        X && Y
        \arrow["{pr_1}"', from=1-2, to=2-1]
        \arrow["{pr_2}", from=1-2, to=2-3]
    \end{tikzcd}\]

    The \textbf{Product Topology} is the coarsest possible topology such that $pr_{1}$ and $pr_{2}$ are both continuous
\end{dfn}

\begin{dfn}[Quotient Topology]{dfn:quotient-top-cont}{}
    For $X$ a topological space and $\sim$ a relation, define a function $q : X \to X /\sim$ where the quotient top is the finest topology such that $q$ is continuous
\end{dfn}


\begin{dfn}[Coarse and Fine Topologies]{dfn:coarse-fine-tops}{}
    For $X$ a topological space and $Y$ a set:
    \begin{itemize}
        \item $f: X \to Y$ means there exists a unique finest topology s.t. $f$ is continuous
        \item $g : Y \to X$ means that there exists a unique coarsest topology s.t. $g$ is continuous
    \end{itemize}
\end{dfn}

\begin{lma}[Hausdorff Coarseness]{lma:hausdorff-coarseness}{}
    $\tau_{1}$ is coarser than $\tau_{2}$ and $\tau_{1}$ is Hausdorff $\implies$ $\tau_{2}$ is Hausdorff
\end{lma}

\newpage
\subsection{Connectedness}

\begin{dfn}[Path Connected]{dfn:path-connected}{}
    A space $X$ is \textbf{path-connected} if for all $x, y\in X$, there exists $\alpha : [0,1] \prightarrow{\text{cts}} X$ s.t. $\alpha(0) = x$ and $\alpha(1) = y$
\end{dfn}

\begin{thm}[]{thm:path-conn-implies-conn}{}
    If a topological space $X$ is path-connected, then it is also connected
\end{thm}

\subsubsection{}
\textbf{Claim}: $\mathbb{R}^{n} \not\cong \mathbb{R}$ for $n > 1$

\newpage
\section{Homotopy Theory}

\subsection{Homotopies}

Recall that for $X$ a topological space, it is path connected iff $\forall x, y\in X$, $\exists \gamma : [0,1] \to X$ such that
\[\gamma(0) = x \quad \text{ and } \quad \gamma(1) = y\]

\textbf{Sameness} $:=$ connectability by a path

\begin{dfn}[Homotopy]{dfn:homotopy}{}
    Let $f, g : X \to Y$ be two continuous maps. A \textbf{homotopy} between them is a continuous map
    \[h : X \times I \to Y\]
    such that $\forall x\in X$, $h(x,0) = f(x)$ and $h(x, 1) = g(x)$
\end{dfn}


\end{document}
