\documentclass[landscape, 8pt]{extarticle}

\usepackage{../../preamble}
\usepackage[separate]{../../rss/thmboxes_v5}
\usepackage{HAlgCustoms}


\begin{document}
	
	%\fontsize{4pt}{5pt} \selectfont %Smallest legible font size when printing with UoE Main Library printers, however some resolution issues for extremely small characters such as subscripts inside fractions. Only use if you *really* need the extra space *cough*HAlg*cough*
	\fontsize{5pt}{6pt} \selectfont
	
	%Reduce horizontal padding in 'tabular' environment, make rows taller for legibility
	\setlength\tabcolsep{4.5pt}
	\setlength\extrarowheight{2pt}
	
	\setlength{\abovedisplayskip}{0pt}
	\setlength{\belowdisplayskip}{0pt}
	\setlength{\abovedisplayshortskip}{0pt}
	\setlength{\belowdisplayshortskip}{0pt}
	
	\begin{multicols*}{4}
		\raggedcolumns 
		%Original template by Leon, modified by Louis. Content by Louis
		\vspace{-5pt}
		\subsection*{HAlg Exam Sheet 2025/26 - What Doesn't Kill You}
		\vspace{-5pt}
		
		
		
		\begin{dfn}[Random Definitions]{dfn:randomstuff}{ }
			\begin{itemize}
				\item pr$_i : X_1 \times ... \times X_n \to X_i, (x_1, ..., x_n) \mapsto x_i$
				\item $\mathcal{P}(X) = \left\{U: U\subseteq X\right\}$ is the \/{power set} of $X$. A subset of $\mathcal{P}$ is a \textbf{system of subsets} of $X$.
				\item 'n-z' means 'non-zero'
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{thm}[Random Theorems]{thm:randoms}{ }
			\begin{itemize}
				\item O-S: For $G$ a group acting on a finite set $X$, $\left|\text{Orb}(x)\right| = \left|G\right|/\left|\text{Stab}(x)\right|$
			\end{itemize}
			
		\end{thm}
		
		
		\begin{rem}[Notation]{rem:noting}{ }
			\begin{itemize}
				\item 'almost all' means 'all but finitely many'
			\end{itemize}
			
		\end{rem}
		
		
	%----------------------------------------------------------------------------------%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% CHAPTER 1 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%----------------------------------------------------------------------------------%
	
	
	{\footnotesize \textbf{\underline{Vector Spaces}}}
		
		
		\begin{dfn}[Systems Of Linear Equations]{dfn:sysoflineqs}{ }
			Let $F$ be a field. We seek to solve the system of $n$ linear equations in $m$ unknowns, described more easily by the extended coefficient matrix.
			\begin{multicols}{2}
			\vspace{-5pt}
			$$
			\begin{array}{ *{4}{cc}c }
				a_{11}x_1 &+& a_{12}x_2 &+& \cdots &+& a_{1m}x_m &=& b_1    \\
				a_{21}x_1 &+& a_{22}x_2 &+& \cdots &+& a_{2m}x_m &=& b_2    \\
				\vdots    & & \vdots    & &        & & \vdots    & & \vdots \\
				a_{n1}x_1 &+& a_{n2}x_2 &+& \cdots &+& a_{nm}x_m &=& b_n
			\end{array} 
			$$
			\columnbreak
			$
			\begin{amatrix}{4}
				a_{11}& a_{12} & \dots & a_{1m} & b_1 \\
				a_{21}& a_{22} & \dots & a_{2m} & b_2 \\
				\vdots& & & \vdots & \vdots \\
				a_{n1}& a_{n2} & \dots & a_{nm} & b_n
			\end{amatrix}
			$
			\end{multicols}
			
			\begin{itemize}
				\item The set $L \subseteq F^m$ of m-tuples satisfying the equations is the \textbf{solution set}.
				\item If the solution set of a linsys of eqns $\ne \emptyset$ we obtain all solutions by adding componentwise an arbitrary solution of the homogenised system to a fixed solution of the system.
			\end{itemize}
			
			
		\end{dfn}
		
		
		\begin{dfn}[Matrices]{dfn:matrices}{ }
			A mapping $A: \left\{1, ..., n\right\} \times \left\{1, ..., m\right\} \to Z$ is called an $n \times m$-matrix with coefficients in $Z$.
			\begin{itemize}
				\item The set of all $n \times m$-matrices with coefficients in $Z$ is denoted Mat$(n \times m; Z) :=$Maps$({1, ..., n} \times {1, ..., m}, Z)$
				\item If $m = n$ we have a \textbf{square matrix} and write Mat$(n, Z)$
				\item For square matrix $A$, its \textbf{trace} $\text{tr}(A)$ is the sum of its diagonal entries
				\item A square matrix $A$ is \textbf{nilpotent} with \textbf{index/degree $k$} if $A^k = 0$ for $k \in \mathbb{N}$. 
				\item $P \in $Mat$(n, \mathbb{R})$ is \textbf{orthogonal} if $P^T P = I_n$, ie $P^T = P^{-1}$.
				\item $P \in $Mat$(n, \mathbb{C})$ is \textbf{unitary} if $\overline{P}^TP = I_n$, ie $\overline{P}^T = P^{-1}$.
				\item $P \in $Mat$(n, \mathbb{C})$ is \textbf{hermitian} if $P = \overline{P}^T$, and the mapping on the standard IPS described by $P$ is self-adjoint.
				\item For $A \in $Mat$(n, \mathbb{C})$, $\exp(A) := \sum_{k = 0}^{\infty} \frac{1}{k!} A^k$. For commutative $A, B$, $\exp(A + B) = \exp(A) \exp(B)$
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{dfn}[Fields And Vector Spaces]{dfn:fieldsnvecs}{ }
			\begin{multicols}{2}
				A \textbf{field} $F$ is a set with functions $+ : F\times F \to F ; (\lambda, \mu) \mapsto \lambda + \mu$ and $\cdot : F \times F \to F; (\lambda, \mu) \mapsto \lambda\mu$ s.t. $(F, +)$ and $(F\backslash\{0\}, \cdot)$ are abelian groups with respective identities $0_F, 1_F$, and $\lambda(\mu + \nu) = \lambda \mu + \lambda\nu \in F$.
				
				\begin{itemize}
					\item $0 \vec{v} = \vec{0}; (-1)\vec{v} = -\vec{v}$
					\item $\forall \lambda \in F$, $\lambda \vec{0} = \vec{0}$
					\item If $\lambda \vec{v} = 0, \lambda = 0$ or $\vec{v} = \vec{0}$
					
				\end{itemize}
			
			
			
				\columnbreak
				A \textbf{vector space $V$ over a field $F$} is a pair made up of an abelian group $V = (V, \dotplus )$ and a mapping $F \times V \to V ; (\lambda, \vec{v} ) \mapsto \lambda\vec{v}$ such that for all $\lambda, \mu \in F$ and $\vec{v}, \vec{w} \in V$ the following hold:
				\begin{align*}
					\lambda(\vec{v} + \vec{w}) &= (\lambda\vec{v}) + (\lambda\vec{w})\\
					(\lambda + \mu)\vec{v} &= (\lambda\vec{v}) + (\mu \vec{v})\\
					\lambda(\mu \vec{v}) &= (\lambda \mu) \vec{v}\\
					1_F \vec{v} &= \vec{v}
				\end{align*}
			
			\end{multicols} 
			
			\begin{itemize}
				\item Let $X$ be a set and $V$ an $F-$VS. Then the set Maps$(X, V)$ of mappings $X \to V$ is an $F-$VS (define $(f+g)(x) = f(x) + g(x)$ and $(\lambda f)(x) = \lambda (f(x))$).
				\item Given $F-$VSs $V_1, ..., V_n$, $V_1 \times ... \times V_n$ is an $F-$VS if we define addition/mult by scalars componentwise. We denote this VS by $V_1 \oplus ... \oplus V_n$ and name it the \textbf{external direct sum}. $F^n = F \oplus ... \oplus F$ $n$ times.
				\item Two VsubSs $V_1, V_2$ of a VS $V$ are \textbf{complementary} if addition defines a bijection: $V_1 \times V_2 \tilde{\to} V$
				
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{dfn}[Vector Subspaces]{dfn:vecsubspc}{ }
			A subset $U$ of a vector space $V$ is a \textbf{vector subspace} if $\vec{0} \in U$ and $\forall \vec{u}, \vec{v} \in U, \lambda \in F$, $\vec{u} + \vec{v}$ and $\lambda \vec{u} \in U$.
			\begin{itemize}
				\item The solution set of a linsys of eqns seen earlier is a vector space
				\item An expression of form $\alpha_1\vec{v}_1 + ... + \alpha_n \vec{v}_n$ with $n$ finite is a \textbf{linear combination} of $\vec{v}_1, ..., \vec{v}_n$.
				\item A subset is a \textbf{hyperplane} if it is a proper subspace and together with one other $\vec{v}$ generates the whole VS.
				\item The intersection of two VSs is itself a VS
			\end{itemize}
			
			Let $T$ be a subset of $V$, a VS over $F$. $\left\langle T\right\rangle$, the \textbf{span of $T$} or \textbf{vector subspace generated by $T$} is the set of all linear combinations of vectors in $T$. (If $T = \emptyset, \left\langle T \right\rangle = \vec{0}$). Note if $v \in \left\langle T\right\rangle, \left\langle T \cup \{v\}\right\rangle = \left\langle T\right\rangle$.\\
			A subset of $V$ is a \textbf{generating/spanning set} if its span is $V$. A VS with a finite generating set is \textbf{finitely generated}.
			\begin{itemize}
				\item Given $V_1, ..., V_n$ VsubSs of $V$, we define the \textbf{sum} $V_1 + ... + V_n$ to be the subspace $\left\langle V_1 \cup ... \cup V_n\right\rangle$ of $V$. If the homo. given by addition$: V_1 \times ... \times V_n$ is also an injection the sum is \textbf{direct} and can be written $V_1 \oplus ... \oplus V_n$.
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{dfn}[Linear Independence]{dfn:linind}{ }
			A subset $L$ of a vector space $V$ is \textbf{linearly independent} if for all $\alpha_1, ..., \alpha_n \in F$ and distinct $\vec{v}_1, ..., \vec{v}_n \in L$, $\alpha_1 \vec{v}_1 + ... + \alpha_n \vec{v}_n = 0$ only if all $\alpha_i = 0$. Otherwise it is \textbf{linearly dependent}.
			\begin{tabular}{|c|c|}
				\hline
				Linearly Independent & Linearly Dependent\\ \hline
				$\emptyset; \{\vec{v}\}, \vec{v}\ne \vec{0}$&$ \{\vec{0}\}$ \\ \hline
			\end{tabular}
			
		\end{dfn}
		
		
		\begin{dfn}[Bases]{dfn:based}{ }
			Let $V$ be a VS over a field $F$. A \textbf{basis} of $V$ is a linearly independent generating set in $V$.
			\begin{itemize}
				\item The family $\left(\vec{v}_i\right)_{1 \le i \le n}$ is a basis of $V$ iff the 'evaluation mapping' $\Phi: F^n \to V; (\alpha_1, ..., \alpha_n) \mapsto (\alpha_1 \vec{v}_1 + ... + \alpha_n \vec{v}_n)$ is a bijection.\\
				\item TFAE for a subset $E$ of a VS $V$:
					\begin{enumerate}[label = (\roman*)]
						\item $E$ is a basis 
						\item $E$ is minimal among generating sets ($\forall \vec{v} \in E$, $E \backslash\vec{v}$ does not generate $V$)
						\item $E$ is maximal among linearly independent subsets ($\forall \vec{v} \in E, E \cup \vec{v}$ is linearly dependent)	
					\end{enumerate}
				\item If $V$ is a finitely generated VS, it has a finite basis
				\item For a set $X$, the set Maps$(X, F)$ equipped with pointwise addition and scalar multiplication is an F-VS. The subset, $F \left\langle X\right\rangle \subseteq$Maps$(X, F)$, of all mappings which send almost all $x \in X$ to zero as a vector subspace, called the \textbf{free vector space on $X$}
				\item TFAE
					\begin{enumerate}[label = (\roman*)]
						\item The family $\left(\vec{v}_i\right)_{i \in I}$ is a basis of $V$.
						\item For all $v \in V$, $\exists$ precisely one family $\left(a_i\right)_{i \in I}$ s.t. almost all $a_i =0$, and $\vec{v} = \sum_{i \in I} a_i \vec{v}_i$.
					\end{enumerate}
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{thm}[FTLA\text{,} Steinitz\text{,} Exchange Lemma]{thm:dfkm}{ }
			Let $V$ be a VS, $L \subset V$ linearly ind., and $E \subseteq V$ generating
			\begin{itemize}
				\item If $V$ is finitely generated then it has a finite basis, and any two bases have the same cardinality
				\item Fundamental Theorem of Linear Algebra: $\left|L\right| \le \left|E\right|$.
				\item Steinitz Exchange Theorem: There is an injection $\phi: L\hookrightarrow E$ s.t. $\{E \backslash \phi(L)\}\cup L$ is also generating.
				\item Exchange Lemma: Steinitz but for single elements.
			\end{itemize}
			
			
		\end{thm}
		
		
		\begin{dfn}[Dimensions Of Vector Spaces]{dfn:dimens}{ }
			Let $V$ be a finitely-generated VS, $L \subset V$ linearly ind., $E \subseteq V$ generating.
			For finitely-generated $V$, its \textbf{dimension}, $\dim V$ is the cardinality of a basis of $V$. For dimension as an $F$-VS write $\dim_F V$. Also:		
			\begin{itemize}
				\item $\left|L\right| \le \dim V$, if $\left|L\right| = \dim V$ then $L$ is a basis
				\item $\left|E\right| \ge \dim V$, if $\left|E\right| = \dim V$ then $E$ is a basis
				\item A proper vector subspace of $V$ has a strictly smaller dimension than $V$
				\item For $U, W$ subspaces of $V$, $\dim(U + W) + \dim (U \cap W) = \dim (U) + \dim (W).$
				\item Given $V_1, ..., V_n$ $F-$VSs, $\dim (V_1 \oplus ... \oplus V_n) = \dim (V_1) + ... + \dim (V_n)$
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{mdfn-s}[Linear Mappings: Homo/Iso/Endo/Automorphisms]{dfn:linmaps}{ }
			Let $V, W$ be VSs over a field $F$ and $B \subset V$ be a basis. A mapping $f: V \to W$ is \textbf{(F-)linear} or a \textbf{homomorphism of $F$-vector spaces} if for all $\vec{v}_1, \vec{v}_2 \in V$ and $\lambda \in F$:\\
			$F(\vec{v}_1 + \vec{v}_2) = f(v_1) + f(v_2)$ and $f(\lambda \vec{v}_1) = \lambda f(\vec{v}_1)$.
			The set of all homomorphisms from $V$ to $W$ is Hom$_F(V, W) =$Hom$(V, W) \subseteq$Maps$(V, W)$
			\begin{itemize}
				\item An \textbf{isomorphism} is a bijective linear mapping. Isomorphisms are invertible, and the inverse mapping is also an isomorphism
				\item An \textbf{endomorphism} is a homomorphism from a VS to itself
				\item An \textbf{automorphism} is an isomorphism from a VS to itself. The automorphisms of a VS $V$ form a subgroup of its permutation group called the \textbf{general linear} or \textbf{automorphism} group, denoted $GL(V) =$Aut$(V) \subseteq Maps^\times(V, V).$
				\item If $f: V \to W$ and $g: U \to V$ are homo.s, $(f \circ g) : U \to W$ is a homo.
				\item The image and preimage of a VS under a linear mapping are VSs.
				\item A homo. can be restricted to a bijection Hom$_F(V, W) \tilde{\to}$Maps$(B, W)$, $f \mapsto \left.f\right|_B$ - ie a linear mapping determines and is determined by the values it takes on a basis.
				\item Every injective homo. $f: V \hookrightarrow W$ has a \textbf{left inverse} $g$ s.t. $g \circ f =$id$_V$
				\item Every surjective homo. $f: V \twoheadrightarrow W$ has a \textbf{right inverse} $g$ s.t. $f \circ g =$id$_W$
				\item For $n \ge 0 \in \mathbb{Z}$, a VS $V$ over $F$ is isomorphic to $F^n$ iff $\dim V = n$.
				\item A point sent to itself by a homo. is a \textbf{fixed point} of the mapping. Given a mapping $f: X \to X$, the set of fixed points is $X^f = \{x \in X : f(x) = x\}$.
				\item Given VS $V$, VsubSs $V_1, ..., V_n$, and linmaps $f_i: V_i \to V$; $f: V_1 \oplus ... \oplus V_n \to V$, $f(v1, ..., v_n) = f_1(v_1) + ... + f_n(v_n)$ is an isomorphism. We then write $V = V_1 \oplus V_2$ and say $V$ is the \textbf{internal direct sum} of $V_1, V_2$.
				\item 
			\end{itemize}
			
		\end{mdfn-s}
		
		
		\begin{dfn}[Image and Kernel]{dfn:imnkern}{ }
			Let $V, W$ be VSs and $f: V \to W$ be a linear mapping.
			The \textbf{image} of $f$ is $(f) = F(V) \subseteq W$. It is a VsubS of $W$ .\\
			The \textbf{kernel} of $f$ is $\ker (f) := f^{-1}(\vec{0}) = \{\vec{v} \in V : f(\vec{v}) = \vec{0}\}$. It is a VsubS of $V$.
			\begin{itemize}
				\item A linear mapping $f$ is injective iff its kernel is $\vec{0}$.
				\item The image of a VsubS under a linear mapping is again a VsubS, and the preimage of a VsubS under a linear mapping is again a VsubS.
			\end{itemize}
			
		\end{dfn}
		
		
		
		\begin{mthm}[Rank-Nullity Theorem]{mthm:ranknull}{ }
			Let $V, W$ be VSs and $f: V \to W$ a linear mapping. \\
			Then $\dim V = \dim (\ker(f)) + \dim($im$(f))$. $\dim($im$(f))$/$\dim (\ker (f))$ are the \textbf{rank/nullity} of $f$.
			
		\end{mthm}
		
		
	%----------------------------------------------------------------------------------------------%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% CHAPTER 2 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%----------------------------------------------------------------------------------------------%
	
	
	{\footnotesize \textbf{\underline{Linear Mappings and Matrices}}}
		
		
		\begin{thm}[Linking Linear Mappings And Matrices]{thm:repmat}{ }
			Let $F$ be a field and $m, n \in \mathbb{N}$. There is a bijection $M$ between the space of linear mappings $F^m \to F^n$ and Mat$(n \times m; F)$ which maps $f \mapsto [f]$. $M(f) := [f]$ is the \textbf{representing matrix} of $f$, whose columns are the images under $f$ of the standard basis elements of $F^m$, ie $[f] = (f(\vec{e}_1)|...|f(\vec{e}_m))$.
			
		\end{thm}
		
		
		\begin{thm}[Composing Mappings vs Matrix Multiplication]{thm:}{ }
			Let $g: F^\ell \to F^m$ and $f:F^m \to F^n$ be linmaps. The representing mat of their composition is the product of their representing mats: $[f \circ g] = [f][g]$.
			\begin{itemize}
				\item For $\ell, m, n \in \mathbb{N}$, $F$ a field, $A \in $Mat$(n\times m; F)$, $B \in$Mat$(m \times \ell ; F)$. Then $AB \in $Mat$(n \times \ell)$ is defined by $(AB)_{ik} = \sum_{j = 1}^{m}A_{ij} B{jk}$.
			\end{itemize}
			
		\end{thm}
		
		
		\begin{dfn}[The General Linear Group]{dfn:glgroup}{ }
			The set of invertible $n\times n$ matrices in a field $F$ form a group under matrix multiplication, called the \textbf{General Linear group of $(n\times n)$ matrices} and written GL$(n; F) :=$Mat$(n; F)^\times$
			
		\end{dfn}
		
		
		
		
		\begin{dfn}[Properties Of Matrices]{dfn:matprops}{ }
			\begin{itemize}
				\item Let $k, \ell, m, n \in \mathbb{N}$, $A, A' \in$Mat$(n\times m; F)$, $B, B' \in $Mat$(m \times \ell; F)$, $C \in$Mat$(\ell \times k; F)$, and $I = I_m$. Then:
				\begin{enumerate}[label=(\roman*)]
					\item $(A + A')B = AB + A'B$
					\item $A(B + B') = AB + AB'$
					\item $(AB)C = A(BC)$
				\end{enumerate}
				\item A square matrix $A$ is \textbf{invertible} if $\exists A^{-1}$ of same size s.t $AA^{-1} = A^{-1}A = I$
				\item An \textbf{elementary} matrix is one which differs from $I$ in at most one entry
				\item Every square matrix with entries in a field can be written as a product of elementary matrices
				\item A matrix whose only non-zero entries lie on the diagonal, and which has first ones along the diagonal, then zeroes, is in \textbf{Smith Normal Form}. For each $A \in $Mat$(n \times m; F)$, $\exists$ invertible $P, Q$ s.t. $PAQ$ is in Smith Normal Form.
				\item The \textbf{rank} of a matrix $A$ is the dimension of the subspace of $F^n$ generated by the columns or rows of $A$ (the $dim$ of each are always equal).\\
				\item A is of \textbf{full rank} if the rank is equal to the number of rows or columns
				\item rk$(A + B) \le$rk$A +$rk$B$, and rk$(AB) \le \min($rk$A,$rk$B)$.
			\end{itemize}
			
			
		\end{dfn}
		
		
		\begin{dfn}[Representing Matrices W.R.T Bases]{dfn:repmatbases}{ }
			Let $F$ be a field, $V, W$ $F-$VSs with ordered bases $\mathcal{A} = (\vec{v}_1, ..., \vec{v}_m)$ and $\mathcal{B} = \vec{w}_1, ..., \vec{w}_n$. \\
			Then to each linmap $F: V \to W$ we associate a \textbf{representing matrix} $_\mathcal{B}[f]_\mathcal{A}$ with entries $a_{ij}$ defined by $f(v_j) = a_{1j}\vec{w}_1 + ... + a_{nj}\vec{w}_n \in W$.
			\begin{itemize}
				\item This produces an isomorphism of VSs $M_\mathcal{B}^\mathcal{A}:$Hom$_F(V, W) \tilde{\to}$Mat$(n\times m; F)$ with $f \mapsto _\mathcal{B}[f]_\mathcal{A}$.
				\item We call $M_\mathcal{B}^\mathcal{A}(f)=_\mathcal{B}[f]_\mathcal{A}$ the \textbf{representing matrix of $f$ with respect to the bases $\mathcal{A}$ and $\mathcal{B}$}. The columns of $_\mathcal{B}[f]_\mathcal{A}$ give the coords of the images of the basis vectors from $\mathcal{A}$ wrt the basis $\mathcal{B}$.
				\item Let $U, V, W$ be finite-dimensional $F-$VSs with ordered bases $\mathcal{A, B, C}$, and $f:U \to V$, $g: V \to W$ be linmaps. Then the representing matrix of $(g \circ f) : U \to W$ is the matrix product of the representing matrices of $f$ and $g$: $_\mathcal{C}[g \circ f]_\mathcal{A} = _\mathcal{C}[g]_{\mathcal{B}\mathcal{B}}[f]_\mathcal{A}$ 
				\item Let $\mathcal{A}$ an ordered basis of a finite-dim VS $V$), the inverse of the 'evaluation mapping' $\Phi_\mathcal{A} : F^m \tilde{\to} V$, $(\alpha_1, ..., \alpha_m)^\text{T} \mapsto \alpha_1 \vec{v}_1 + ... + \alpha_m \vec{v}_m$ is given by $\vec{v} \mapsto _\mathcal{A}[\vec{v}]$. $_\mathcal{A}[\vec{v}]$ is the \textbf{representation of $\vec{v}$ w.r.t. the basis $\mathcal{A}$}.
				\item Let $\mathcal{A, B}$ be ordered bases of finite-dim VSs $V, W$ and $f: V \to W$ be a linmap. Then for $\vec{v} \in V$, $_\mathcal{B}[f(\vec{v})] = _\mathcal{B}[f]_{\mathcal \mathcal{A}}[\vec{v}]$
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{dfn}[Change Of Matrix By Change Of Basis]{dfn:chngofbasismat}{ }
			Let $\mathcal{A} = (\vec{v}_1, ..., \vec{v}_n)$ and $\mathcal{B} = (\vec{w}_1, ..., \vec{w}_n)$ both be ordered bases of an $F-$VS $V$. Then the matrix representing the identity map w.r.t. these bases, $_\mathcal{B}[\text{id}_V]_\mathcal{A}$ is called a \textbf{change of basis matrix}. By definition its entries are given by the equalities $\vec{v}_j = \sum_{i=1}^{n} a_{ij}\vec{w}_i$.
			\begin{itemize}
				\item Let $V, W$ be finite-dim $F-$VSs and $f: V \to W$ be a linmap. Suppose $\mathcal{A, A'}$ and $\mathcal{B, B'}$ are ordered bases of $V$ and $W$ respectively. Then $_\mathcal{B'}[f]_\mathcal{A} = _\mathcal{B'}[\text{id}_W]_{\mathcal{B} \mathcal{B}}[f]_{\mathcal{A} \mathcal{A}}[\text{id}_V]_\mathcal{A'}$.
				\item For $\mathcal{A, A'}$ ordered bases of finite-dim VS $V$, and $f$ an endo. of $V$; $_\mathcal{A'}[f]_\mathcal{A'} = _\mathcal{A}[f]_{\mathcal{A'} \mathcal{A}}^{-1}[f]_{\mathcal{A} \mathcal{A}}[\text{id}_V]_\mathcal{A'}$
				\item $N = _\mathcal{B}[f]_\mathcal{B}$, $M = _\mathcal{A}[f]_\mathcal{A}$ are \textbf{similar} if $N = T^{-1} M T$ ($T := _\mathcal{A}[\text{id}_V]_\mathcal{B}$).
				\item Let $f: V \to W$ be a linmap between finite-dim $F-$VSs. There exists ordered bases $\mathcal{A}$ of $V$ and $\mathcal{B}$ of $W$ s.t. $_\mathcal{B}[f]_\mathcal{A}$ is in Smith Normal Form
			\end{itemize}
			
		\end{dfn}
		
		
	%----------------------------------------------------------------------------------%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% CHAPTER 3 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%----------------------------------------------------------------------------------%
	
	
	{\footnotesize \textbf{\underline{Rings And Modules}}}
		
		
		\begin{mdfn}[A Ring]{mdfn:ring}{ }
			A \textbf{ring} is a set $R$ equipped with two operations $+, \cdot$ such that:\\
			\begin{enumerate}[label=(\roman*)]
				\item $(R, +)$ is an abelian group
				\item $(R, \cdot)$ is a monoid (satisfies associativity and identity)
				\item $a \cdot (b + c) = (a \cdot b) + (a\cdot c)$ and $(a + b) \cdot c = (a \cdot c) + (b \cdot c)$.
			\end{enumerate}
			The element $1 \in R$ is uniquely defined as the identity of $(R, \cdot)$.
			\begin{itemize}
				\item A ring in which $\cdot$ is commutative is a \textbf{commutative ring}.
				\item The \textbf{null/zero ring} is the ring where $R$ is a single element set.
				\item The integers modulo $m$, or $\mathbb{Z}/m\mathbb{Z}$ is a ring, with congruence classes as elements.
				\item A field is a n-z commutative ring $F$ where every n-z $a \in F$ has inverse $a^{-1} \in F$, ie $a \cdot a^{-1} = a^{-1} \cdot a = 1$
				\item For $m \in \mathbb{N}$, $\mathbb{Z}/m\mathbb{Z}$ is a field iff $m$ is prime(
				\item For $a, b \in R$, $0a = 0 = a0$; $(-a)b = -(ab) = a(-b)$; and $(-a)(-b) = ab$.
				\item For $m \in \mathbb{Z}$, the \textbf{$m$-th multiple ma of $a \in$ abelian $R$} is $ma := a+...+a$ m times if $m>0$, $0a = 0$. If $m < 0$, $(-m)(a) = -(ma)$.
				\item For $m, n \in \mathbb{Z}$, $a, b \in R$: $m(a+b) = ma + mb$; $(m + n)a = ma + na$; $m(na) = (mn)a$; and $m(ab) = (ma)b = a(mb)$.
			\end{itemize}
			
		\end{mdfn}
		
		
		\begin{mdfn}[Units\text{,} Zero-Divisors\text{,} Integral Domains]{mdfn:unitsnzerodivsnintdoms}{ }
			Let $R$ be a ring, and $a \in R$.\\
			$a$ is a \textbf{unit} if it is invertible in $R$ ($\exists a^{-1} \in R$ s.t. $aa^{-1} = a^{-1}a = 1$). \\
			A n-z $a$ is a \textbf{zero-divisor} if $\exists$ n-z $b \in R$ s.t. either $ab = 0$ or $ba=0$.\\
			An \textbf{integral domain} is a n-z commutative ring with no zero-divisors.
			\begin{itemize}
				\item The set of units of a ring, $R^\times$, form a group under multiplication.
				\item For $m \in \mathbb{N}$, $\mathbb{Z}/m\mathbb{Z}$ is an integral domain iff $m$ is prime
				\item Every finite integral domain is a field
				\item Let $a, b, c \in R$, an int-dom. If $ab = ac$ and $a \ne 0$, then $b = c$.
			\end{itemize}
			
		\end{mdfn}
		
		
		\begin{dfn}[Polynomials]{dfn:polynoms}{ }
			Let $R$ be a ring. A \textbf{polynomial over $R$} is an expression of the form $P = a_m x^m + ... + a_2 x^2 + a_1 x + a_0$ for $0 \le m \in \mathbb{Z}$ and $a_i \in R$.
			\begin{itemize}
				\item The set of all polynomials over $R$ is denoted $R[x]$.
				\item When the leading coefficient is one, the polynomial is \textbf{monic}
				\item Defined with the usual addition/multiplication rules for polynomials, $R[x]$ becomes a ring, called the \textbf{ring of polynomials with coefficients in $R$}. The zero/identity of $R$ are the zero/identity of $R[x]$.
				\item If $R$ is commutative/an integral domain/has no zero-divisors, so is or does $R[x]$
				\item If $R$ has no zero-divisors, $\deg(PQ) = \deg(P) + \deg(Q)$ for n-z $P, Q \in R[x]$.
				\item Division and Remainder Thm: Let $R$ be an int-dom and $P, Q \in R[x]$ with $Q$ monic. Then $\exists$ unique $A, B \in R[x]$ s.t. $P = AQ + B$; and $\deg (B) < \deg (Q)$ or $B=0$
				\item For $R$ a commutative ring $P \in R[x]$ can be \textbf{evaluated} at $\lambda \in R$ to give $P(\lambda)$. In this way we have a mapping $R[x] \to $Maps$(R, R)$
				\item $\lambda \in R$ is a root of $P$ if $P(\lambda) = 0$, and iff $(x - \lambda)$ divides $P(x)$.
				\item For $R$ an int-dom, n-z $P \in R[x] \backslash \{0\}$ has at most $\deg (P)$ roots in $R$.
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{dfn}[Algebraic Closure]{dfn:algclos}{ }
			A field $F$ is \textbf{algebraically closed} if each non-constant polynom $P \in F[x] \backslash F$ with coefficients in $F$ has a root in $F$.\\
			\begin{itemize}
				\item If $F$ is algebraically closed, every n-z $P \in F[x] \backslash \{0\}$ decomposes uniquely (up to reordering) into linear factors: $P = c(x-\lambda_1)...(x-\lambda_n)$ with $n \ge 0, c \in F^\times, \lambda_1, ..., \lambda_n \in F$.
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{mdfn}[Ring Homomorphisms]{mdfn:ringhomos}{ }
			Let $R, S$ be rings. A mapping $f:R \to S$ is a \textbf{ring homomorphism} if for all $x, y \in R$: $f(x + y) = f(x) + f(y)$ and $f(xy) = f(x)f(y)$.
			\begin{itemize}
				\item The \textbf{zero ring homomorphism} is $f: R \to S; r \mapsto 0$.
				\item \textbf{Isomorphisms} etc are akin to linear mappings
				\item If $S$ has no zero-divs then either $f(1_R) = 0_S$ and $f$ is the zero homo., or $f(1_R) = 1_S$.
				\item The composition of two homo.s is a homo., and the inverse of an iso. is an iso.
				\item If $f: R \to S$ a ring homo., then $\forall x, y \in R$ and $m \in \mathbb{Z}$: $f(0_R) = 0_S$; $f(-x) = -f(x)$; $f(x-y) = f(x) - f(y)$; $f(mx) = mf(x)$.
			\end{itemize}
			
		\end{mdfn}
		
		
		\begin{mdfn}[Ideals]{mdfn:ideals}{ }
			A subset $I$ of a ring $R$ is an \textbf{ideal}, written $I \trianglelefteq R$, if it is nonempty; closed under subtraction; and closed under multiplication by elements of $R$.
			\begin{itemize}
				\item For $R$ a comring and $T \subset R$, the \textbf{ideal of $R$ generated by $T$} is the set $_R \left\langle T\right\rangle = \{r_1t_1 + ... + r_mt_m: t_1, ..., t_m \in T, r_1, ..., r_m \in R\}$ together with $0_R$ if $T = \emptyset$. If $T$ is finite often write $_R \left\langle t_1, ..., t_n\right\rangle$.
				\item For $R$ a comring and $T \subseteq R$, $_R \left\langle T\right\rangle$ is the smallest ideal of $R$ containing $T$.
				\item For $R$ a comring, an ideal $I$ is a \textbf{principal ideal} if $I = \left\langle t\right\rangle$ for some $t \in R$. In a comring, $\{0\}$ and $R$ are always principal ideals, generated by $0$ and $1_R$.
			\end{itemize}
			
		\end{mdfn}
		
		
		\begin{dfn}[The Kernel Of A Ring Homomorphisms]{dfn:ringker}{ }
			Let $R, S$ be rings with zero elements $0_R, 0_S$ and $f: R \to S$ be a ring homo. Then the \textbf{kernel} of $f$ is $\ker f - \left\{r \in R : f(r) = 0_S\right\}$.
			\begin{itemize}
				\item $\ker f$ is an ideal of $R$.
				\item $f$ is injective iff $\ker f = \{0\}$.
				\item The intersection of any collection of ideals is an ideal
				\item For $I, J$ ideals of $R$; $I + J := \{a + b : a \in I, b \in J\}$ is an ideal
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{dfn}[Subrings]{dfn:subrings}{ }
			Let $R$ be a ring. A subset $R'$ of $R$ is a subring of $R$ if $R'$ is a ring under the same operations $+/\cdot$ defined in $R$.\\
			Test For A Subring: A subset $R'$ of a ring $R$ is a subring if and only if:
			\begin{enumerate}[label=(\roman*)]
				\item $R'$ has a multiplicative identity ($1_{R'}$ is not necessarily equal to $1_R$!)
				\item $R'$ is closed under subtraction ($a, b \in R \implies a-b \in R$)
				\item $R'$ is closed under multiplication
			\end{enumerate}
			The intersections of two subrings of $R$ is not necessarily a subring of $R$.
			Let $R, S$ be rings and $f: R \to S$ a ring homomorphism. Then:
			\begin{itemize}
				\item If $R'$ is a subring of $R$ then $f(R')$ is a subring of $S$. Particularly, im$f$ is a subring of $S$.
				\item Let $f(1_R) = 1_S$. Then if $x$ is a unit in $R$, $f(x)$ is a unit in $S$ and $(f(x))^{-1} = f(x^{-1})$. In this case $f$ restricts to a group homomorphism $\left.f\right|_{R^\times} : R^\times \to S^\times$.
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{dfn}[Equivalence]{dfn:equiv}{ }
			A \textbf{relation} $R$ on a set $X$ is a subset $R \subseteq X \times X$. Only in this context, we write $(x, y) \in R$ as $xRy$.\\
			$R$ is an \textbf{equivalence relation on $X$} when for all $x, y, z, \in X$:
			\begin{enumerate}[label=(\roman*)]
				\item Reflexivity: $xRx$
				\item Symmetry: $xRy$ if and only if $yRx$
				\item Transitivity: ($xRy$ and $yRz$) $\implies xRz$
			\end{enumerate}
			\begin{itemize}
			\item[*] Let $~$ be an equivrel on a set $X$. For $x \in X$ the set $E(x) := \left\{z \in X : z~ x\right\}$ is the \textbf{equivalence class of $x$}. $E \subseteq X$ is an \textbf{equivalence class for $~$} if $\exists x \in X$ s.t. $E = E(x)$.
			\item  An element of an equivalence class is a \textbf{representative} of the class. 
			\item $Z \subseteq X$ is a \textbf{system of representatives for $~$} if it contains precisely one element from each equivalence class of $~$. 
			\item For $x, y \in X$ TFAE: $x ~ y$; $E(x) = E(y)$; $E(x) \cap E(y) \ne \emptyset$
			\item[*] Let $~$ be an equivrel on a set $X$. We denote the \textbf{set of equivalence classes of $X$}, a subset of $\mathcal{P}(X)$, as $(X/~) := \left\{E(x) : x \in X\right\}$.
			\item There is a surjective canonical mapping can$: X \twoheadrightarrow (X/~), x \mapsto E(x)$.
			\item Let $~$ be an equivrel on a set $X$. If $f: X \to Z$ is a mapping s.t. $x ~ y \implies f(x) = f(y)$; there is a unique mapping $\overline{f} : (X/~) \to Z$, $\overline{f}(E(x)) := f(x)$, and $f = \overline{f} \circ$can.
			\item Often when defining a map $g: (X/~) \to Z$, we define $f: X \to Z$ and say $g$ is \textbf{well-defined} to indicate that $x ~ y \implies f(x) = f(y)$ as this guarantees the existence of $g = \overline{f}$.
			\end{itemize}			
			
		\end{dfn}
		
		
		\begin{dfn}[Cosets and Factor Rings]{dfn:cosetsnfacrings}{ }
			Let $I \trianglelefteq R$ be an ideal in a ring $R$. The set $x + I := \{x + i : i \in I\} \subseteq R$ is a \textbf{coset of $I$ in $R$} or the \textbf{coset of $X$ with respect to $I$ in $R$}.\\
			Let $R$ be a ring, $I \trianglelefteq R$ an ideal, and $~$ the equivrel defined by $x ~ y \iff x-y \in I$. Then $R/I$, the \textbf{factor ring} or \textbf{quotient of $R$ by $I$}, is the set $(R/~)$ of cosets of $I$ in $R$.\\
			 Let $R$ be a ring and $I \trianglelefteq R$ an ideal. Then:
			\begin{itemize}
				\item $R/I$ is a ring, with addition defined by $(x+I) + (y+I) = (x+y)+I$ and multiplication by $(x+I)(y+I) = xy + I$ for all $x, y \in R$.
				\item If $R$ is commutative so is $R/I$
				\item $R/I$ is a nonzero ring if and only if $I \ne R$
				\item If $I$ is proper, and $r \in R^\times$; $r+I \in (R/I)^\times$ with $(r+I)^{-1} = r^{-1}+I$.
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{thm}[The Universal Property Of Factor Rings]{thm:univpropfacrings}{ }
			Let $R$ be a ring and $I \trianglelefteq R$. Then:
			\sbsadaptr{
				\begin{enumerate}
					\item The mapping can$:R \to R/I, r \mapsto r + I$ is a surjective ring homo. with kernel $I$.
					\item If $f: R \to S$ is a ring homo. with $f(I) = \{0_S\}$, ie $I \subseteq \ker f$, there is a unique ring homo. $\overline{f} : R/I \to S$ s.t. $f = \overline{f} \circ $can.
				\end{enumerate}
			}
			{
			% https://q.uiver.app/#q=WzAsMyxbMCwwLCJSIl0sWzIsMCwiUi9JIl0sWzIsMiwiUyJdLFswLDEsIlxcdGV4dHtjYW59Il0sWzAsMiwiZiIsMl0sWzEsMiwiXFxvdmVybGluZXtmfSIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImRhc2hlZCJ9fX1dXQ==
			\begin{tikzcd}[ampersand replacement=\&,cramped,sep=scriptsize]
				R \&\& {R/I} \\
				\\
				\&\& S
				\arrow["{\text{can}}", from=1-1, to=1-3]
				\arrow["f"', from=1-1, to=3-3]
				\arrow["{\overline{f}}", dashed, from=1-3, to=3-3]
			\end{tikzcd}
			}
			
		\end{thm}
		
		
		\begin{mthm}[The First Isomorphism Theorem For Rings]{mthm:fitforrings}{ }
			Let $R, S$ be rings. Then every ring homo. $f: R \to S$ induces a ring iso.:
			$$
			\overline{f} : R/\ker f \tilde{\to} \text{im}f
			$$
			
		\end{mthm}
		
		
		
		\begin{dfn}[Modules]{dfn:modules}{ }
			A \textbf{(left) module $M$ over a ring $R$} or an \textbf{$R$-module} is a pair consisting of an abelian group $M = (M, \dotplus)$ and a mapping $R \times M \to M; (r, a) \mapsto ra$ such that for all $r, s \in R$ and $a, b \in M$ the following hold:
			\begin{multicols}{2}
				\begin{itemize}
					\item $r(a \dotplus b) = (ra) \dotplus (rb)$
					\item $(r + s)a = (ra) \dotplus (sa)$
				\end{itemize}
				\columnbreak
				\begin{itemize}
					\item $r(sa) = (rs)a$
					\item $1_Ra = a$
				\end{itemize}
			\end{multicols}
			Let $R$ be a ring and $M$ an $R-$module. Then for all $r \in R, a \in M$: $0_R a = 0_M$; $r0_M = 0_M$; $(-r)a = r(-a) = -(ra)$
			
		\end{dfn}
		
		
		\begin{dfn}[Module Homomorphisms/Kernel/Image/etc]{dfn:modulestuff}{ }
			Let $R$ be a ring and $M, N$ be $R-$modules.\\
			A mapping $f: M \to N$ is an \textbf{($R$-)homomorphism} if for all $a, b \in M$ and $r \in R$: $f(a+b) = f(a) + f(b)$, and $f(ra) = rf(a)$.\\
			The \textbf{kernel} of $f$ is $\ker f - \{a \in M : f(a) = 0_N\} \subseteq M$ and the \textbf{image} of $f$ is im$f = \{f(a) : a \in M\} \subseteq N$.\\
			If $f$ is a bijection then it is an \textbf{($R$-module) isomorphism} with $M \cong N$.
			\begin{itemize}
				\item The composition of two homo.s is a homo.
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{dfn}[Submodules]{dfn:submods}{ }
			A non-empty subset $M'$ of an $R-$module $M$ is a \textbf{submodule} if $M'$ is an $R-$module with respect to the operations of the $R-$module $M$ restricted to $M'$.\\
			Test For A Submodule: Let $R$ be a ring and $M$ be an $R-$module. $M' \subseteq M$ is a submodule iff:
			\begin{enumerate}[label=(\roman*)]
				\item $0_M \in M'$
				\item $a, b \in M' \implies a-b \in M'$
				\item $r \in R$ and $a \in M' \implies ra \in M'$
			\end{enumerate}
			Let $R$ be a ring, $M, N$ be $R-$modules, $f : M \to N$ be an $R-$homo.. Then:
			\begin{itemize}
				\item $\ker f$ is a submodule of $M$ and im$f$ is a submodule of $N$.
				\item $f$ is injective if and only if $\ker f = {0_M}$
			\end{itemize}
			
			Let $R$ be a ring, $M$ an $R-$module, and $T \subseteq M$. Then the \textbf{submodule of $M$ generated by $T$} is the set $_R \left\langle T\right\rangle = \{r_1t_1 + ... + r_mt_m: t_1, ..., t_m \in T, r_1, ..., r_m \in R\}$ together with the zero element if $T = \emptyset$.
			Submodules have several properties akin to subspaces/ideals:
			\begin{itemize}
				\item M is \textbf{finitely generated} if it is generated by a finite set
				\item M is called \textbf{cyclic} if it is generated by a singleton
				\item $_R \left\langle T\right\rangle$ is the smallest submodule of $M$ containing $T$, for $T \subseteq M$.
				\item The intersection of any collection of submodules of $M$ is a submodule of $M$
				\item For $M_1, M_2$ submodules of $M$, $M_1 + M_2 := \{a + b : a \in M_1, b \in M_2\}$ is a submodule of $M$.
				\item 
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{dfn}[Factor Modules]{dfn:facmods}{ }
			Let $R$ be a ring, $M$ an $R-$module, and $N$ a submodule of $M$. For each $a \in M$, the \textbf{coset of $a$ with respect to $N$ in $M$} is : $a + N = \{a+b : b \in N\}$.\\
			It is a coset of $N$ in the abelian group $M$, so is an equiv class for the equivrel $a ~ b \iff a-b \in N$.\\
			$M/N$, the \textbf{factor/quotient of $M$ by $N$} to be the set $(M/~)$ of all cosets of $N$ in $M$. This becomes an $R-$module by defining: $(a+N) \dotplus (b+N) := (a+b) + N$ and $r(a+N) = ra + N$. for all $a, b \in M, r \in R$.\\
			The $R-$module $M/N$ is the \textbf{factor module} of $M$ by the submodule $N$.
			
		\end{dfn}
		
		
		\begin{thm}[Universal Property Of Factor Modules]{thm:univpropfacmods}{ }
			Let $R$ be a ring, $L, M$ be $R-$modules, and $N$ a submodule of $M$:
			\sbsadaptr{
			\begin{enumerate}
				\item The mapping can$: M \to M/N$, $a \mapsto a+N$ for all $a \in M$ is a surjective $R-$homo. with kernel $N$.
				\item If $f: M \to L$ is an $R-$homo with $f(N) = \{0_L\}$ so that $N \subseteq \ker f$, there exists unique homo. $\overline{f}: M/N \to L$ such that $f = \overline{f} \circ $can. 
			\end{enumerate}}
			{
				% https://q.uiver.app/#q=WzAsMyxbMCwwLCJNIl0sWzIsMCwiTS9OIl0sWzIsMiwiTCJdLFswLDEsIlxcdGV4dHtjYW59Il0sWzAsMiwiZiIsMl0sWzEsMiwiXFxvdmVybGluZXtmfSIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImRhc2hlZCJ9fX1dXQ==
				\begin{tikzcd}[ampersand replacement=\&,cramped, sep=scriptsize]
					M \&\& {M/N} \\
					\\
					\&\& L
					\arrow["{\text{can}}", from=1-1, to=1-3]
					\arrow["f"', from=1-1, to=3-3]
					\arrow["{\overline{f}}", dashed, from=1-3, to=3-3]
				\end{tikzcd}
			}
			
		\end{thm}
		
		
		\begin{mthm}[First Isomorphism Theorem For Modules]{thm:firstisothmformods}{ }
			Let $R$ be a ring and $M, N$ be $R-$modules. Every $R-$homo. $f: M \to N$ induces an $R-$iso.:\\
			$\overline{f}: M/\ker f \tilde{\to}$im$f$
			\begin{itemize}
				\item 2$^{nd}$ Theorem: For $K, L$ sub-mods of $M$; $L$ is a submod of $K + L = \{b + c : b \in K, c \in L\}$ and $K \cap L$ is a submod of $K$. Also, $\frac{K + L}{L} \cong \frac{K}{K \cap L}$
				\item 3$^{rd}$ Theorem: For $K, L$ sub-mods of $M$, with $L \subseteq K$. Then $K/L$ is a submod of $M/L$ and $\frac{M/L}{K/L} \cong M/K$.
			\end{itemize}
			
			
		\end{mthm}
		
		
		
	%----------------------------------------------------------------------------------%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% CHAPTER 4 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%----------------------------------------------------------------------------------%
	
	
	{\footnotesize \textbf{\underline{Determinants And Eigenvalues}}}
		
		
		
		\begin{dfn}[Permutations And Their Signs]{dfn:permsnsigns}{ }
			The group of all permutations of the set $\{1, ..., n\}$ to itself, denoted $\mathfrak{S}_n$, is the \textbf{$n-$th symmetric group}. It is a group under composition with order $n!$
			\begin{itemize}
				\item A \textbf{transposition} is a permutation which swaps two elements and leaves the others unchanged
				\item An \textbf{inversion} of a $\sigma \in \mathfrak{S}_n$ is a pair $(i, j)$ s.t. $1 \le i < j \le n$ and $\sigma (i) > \sigma (j)$.\\
				\item The number of inversions of a $\sigma $ is the \textbf{length} of $\sigma $, $\ell(\sigma )$
				\item The \textbf{sign} of $\sigma$ is the parity of the number of inversions of $\sigma$ - sgn$(\sigma) = (-1)^{\ell(\sigma_)}$. 
				\item If sgn$(\sigma) = 1$, $\sigma$ is an \textbf{even permutation}. If it $= -1$, $\sigma$ is an \textbf{odd permutation}
				\item $\text{sgn}(\sigma \tau) = \text{sgn}(\sigma)\text{sgn}(\tau)$
				\item The set of even perms. in $\mathfrak{S}_n$ forms a subgroup of $\mathfrak{S}_n$ called the \textbf{alternating group} and denoted $A_n$.
				
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{dfn}[Cofactors And Adjugate Matrices]{dfn:cofacsnadjs}{ }
			Let $A \in $Mat$(n; R)$ for a comring $R$, $n, i, j \in \mathbb{N}$ s.t. $1 \le i, j \le n$. Then: 
			\begin{itemize}
				\item The $(i, j)$ \textbf{cofactor} of $A$ is $C_{ij} = (-1)^{i + j}\det(A\left\langle i, j\right\rangle)$ where $A \left\langle i, j\right\rangle$ is the matrix obtained by deleting the $i^{th}$ row and $j^{th}$ column of $A$.
				\item The \textbf{adjugate matrix} $\text{adj}(A)$ is the $(n\times n)$ matrix whose entries are adj$(A)_{ij} = C_{ji}$ \textit{(note order swap of $i$ and $j$!)}.
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{mdfn}[The Determinant]{mdfn:det}{ }
			Let $R$ be a comring and $n \in \mathbb{N}$. The \textbf{determinant} is a mapping $\det : $Mat$(n; R) \to R$ given by: $\det(A) = \sum_{\sigma \in \mathfrak{S}_n} \text{sgn}(\sigma)a_{1 \sigma(1)} ... a_{n \sigma(n)}$. This is called the \textbf{Leibniz formula}. For $n=0$ $\det(A) = 1$. \\
			Alternative characterisations are:
			\begin{itemize}
				\item Let $F$ be a field. The mapping $\det : $Mat$(n; F) \to F$ is the unique alternating multilinear form on $n-$tuples of column vectors with values in $F$ that takes the value $1_F$ on the identity matrix.
				\item Laplace Expansions: Let $R$ be a comring and $A = (a_{ij}) \in $Mat$(n; R)$. For fixed $i, j$ the \textbf{$i^{th}$ row/$j^{th}$ column expansion of the determinant} is $\det(A) = \sum_{j=1}^{n}a_{ij}C_{ij}$/$\sum_{i=1}^{n}a_{ij}C_{ij}$.
			\end{itemize}
			
		\end{mdfn}
		
		
		\begin{thm}[Calculations With The Determinant]{thm:detcalcs}{ }
			\begin{itemize}
				\item Let $R$ be a comring and $A, B \in $Mat$(n; R)$. Then $\det(AB) = \det(A) \det(B)$; $\det(A^\text{T}) = \det(A)$
				\item For $A, B$ same-size square matrices with entries in a field $F$: $\det(A) \ne 0$ iff the $A$ is invertible; if $A$ is invertible then $\det(a^{-1}) = \det(A)^{-1}$ and $\det(A^{-1}BA) = \det(B)$ (ie, the $\det$ of an endomorphism is independent of the choice of basis used to calculate it)
				\item When performing Gaussian Elimination, row addition does not change the determinant, and row swaps change the sign only
			\end{itemize}
			
		\end{thm}
		
		
		\begin{dfn}[Bi/Multilinear Forms]{dfn:bilmultilforms}{ }
			Let $U, V, W$ be $F-$VSs. A \textbf{bilinear form on $U \times V$ with values in $W$} is a mapping $H: U\times V \to W$ which is linear in both entries, ie for all $u_1, u_2 \in U$ and $v_1, v_2 \in V$:
			\begin{multicols}{2}
				\begin{itemize}
					\item $H(u_1 + u_2, v_1) = H(u_1, v_1) + H(u_2, v_1)$
					\item $H(u_1, v_1 + v_2) = H(u_1, v_1) + H(u_1, v_2)$
				\end{itemize}
				\columnbreak
				\begin{itemize}
					\item $H(\lambda u_1, v_1) = \lambda H(u_1, v_1)$
					\item $H(u_1, \lambda v_1) = \lambda H(u_1, v_1)$
				\end{itemize}
			\end{multicols}
			For $U = V$, a bilinear form is \textbf{symmetric} if for all $u, v \in U$, $H(u, v) = H(v_u)$; and it is \textbf{alternating/antisymmetric} if for all $u \in U$, $H(u, u) = 0$
			\begin{itemize}
				\item If $H$ is antisymmetric, $H(u, v) = -H(v, u)$
			\end{itemize}
			
			\textbf{Multilinear forms} are just an extension of bilinear forms to higher dimensions - $H$ must be linear in each argument.
			\begin{itemize}
				\item A multilinear form is \textbf{alternating} if $\exists i \ne j$ with $v_i = v_j \implies H(v_1, ..., v_i, ..., v_j, ..., v_n) = 0$ and if so, has the property $H(v_1, ..., v_i, ..., v_j, ..., v_n) = -H(v_1, ..., v_j, ..., v_i, ..., v_n)$ for all $v_1, ..., v_n \in V$.
				\item Thus for any $\sigma \in \mathfrak{S}_n$, $H(v_{\sigma(1)}, ..., v_{\sigma(n)}) =\text{sgn}(\sigma)H(v_1, ..., v_n)$
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{thm}[Cramer's Rule]{thm:cramer}{ }
			Let $R$ be a comring and $A \in $Mat$(n; R)$. Then $A \cdot $adj$A = (\det(A)) I_n$
			\begin{itemize}
				\item $A$ is invertible iff $\det(A)$ is a unit in $R$, ie $\det(A) \in R^\times$
			\end{itemize}
			
		\end{thm}
		
		
		\begin{thm}[Jacobi's Formula]{thm:jacobi}{ }
			Let $A = (a_{ij})$ where the coeffs $a_{ij} = a_{ij}(t)$ are functions of $t$. Then $\frac{d}{dt}\det(A) = $tr adj$A \frac{dA}{dt}$
			
		\end{thm}
		
		
		\begin{mdfn}[Eigenvalues and Eigenvectors]{mdfn:eigenstuff}{ }
			Let $f: V \to V$ be an endomorphism of an $F-$VS $V$. $\lambda \in F$ is an \textbf{eigenvalue of $F$} iff there exists n-z $\vec{v} \in V$ s.t. $f(\vec{v}) = \lambda \vec{v}$.\\
			Each such $\vec{v} \in V$ is called an \textbf{eigenvector of $f$ with eigenvalue $\lambda$}.\\
			For any $\lambda \in F$, the \textbf{eigenspace of $f$ with eigenvalue $\lambda$} is $E(\lambda, f) = \{\vec{v} \in V : f(\vec{v}) = \lambda\vec{v}\}$.
			\begin{itemize}
				\item Each endomorphism of a n-z finite-dim VS over an algebraically closed field has an eigenvalue
				\item If $\vec{v}_1, ..., \vec{v}_n$ are eigenvecs of $f$ with distinct eigenvals $\lambda_1, ..., \lambda_n$, then $\vec{v}_1, ..., \vec{v}_n$ are linearly independent.
			\end{itemize}
			
		\end{mdfn}
		
		
		\begin{dfn}[The Characteristic Polynomial]{dfn:charpoly}{ }
			Let $R$ be a comring and $A \in $Mat$(n; R)$. The polynomial $\chi_A(x) := \det(xI_n - A) \in R[x]$ is called the \textbf{characteristic polynomial of $A$}.
			\begin{itemize}
				\item For $F$ a field and $A \in $Mat$(n; F)$. The eigenvalues of the linmap $A: F^n \to F^n$ are exactly the roots of $\chi_A$.
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{dfn}[Conjugacy]{dfn:conj}{ }
			For a comring $R$, $A, B \in $Mat$(n; R)$, are \textbf{conjugate} if $B = P^{-1}AP$ for invertible $P \in $GL$(n; R)$. 
			\begin{itemize}
				\item Conjugacy is an equivalence relation on Mat$(n; R)$
				\item The characteristic polynomials of conjugate matrices are the same
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{dfn}[Triangularisability]{dfn:triang}{ }
			Let $f: V \to V$ be an endomorphism of a finite-dim $F-$VS $V$. The following two statements are equivalent, and if they hold, $f$ is \textbf{triangularisable}:
			\begin{enumerate}[label=(\arabic*)]
				\item The characteristic polynom of $f$, $\chi_f (x)$, decomposes into linear factors in $F[x]$.
				\item $V$ has an ordered basis $\mathcal{B} = (\vec{v}_1, \dots, \vec{v}_n)$ s.t. $f(\vec{v}_1) = a_{11}\vec{v}_1, f(\vec{v}_2) = a_{12}\vec{v}_1 + a_{22} \vec{v}_2, \dots, f(\vec{v}_n) = a_{1n}\vec{v}_1 + ... + a_{nn}\vec{v}_n \in V$. Equivalently, the matrix $_\mathcal{B}[f]_\mathcal{B}$ is upper triangular.
			\end{enumerate}
			\begin{itemize}
				\item An endomorphism $A : F^n \to F^n$ is triangularisable iff $A$ is conjugate to an upper triangular matrix $B$. Thus any endomorphism of a finite-dim VS over an algebraically closed field is triangularisable.
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{dfn}[Diagonalisability]{dfn:diag}{ }
			An endo. $f: V \to V$ of an $F-$VS is \textbf{diagonalisable} iff there exists a basis of $V$ consisting of eigenvectors of $f$. If $V$ is finite-dim, then this is equivalent to: $\exists$ ordered basis $\mathcal{B} = \{\vec{v}_1, ..., \vec{v}_n\}$ such that $_\mathcal{B}[f]_\mathcal{B} = $diag$(\lambda_1, ..., \lambda_n)$. Here $f(\vec{v}_i) = \lambda_i \vec{v}_i$.\\
			A square matrix $A \in $Mat$(n; F)$ is \textbf{diagonalisable} iff the linear mapping $F^n \to F^n$ given by left multiplication by $A$ is diagonalisable. That is, $A$ is conjugate to a diagonal matrix ($\exists P \in $GL$(n; F)$ s.t. $P^{-1}AP = $diag$(\lambda_1, ..., \lambda_n)$.) In this case the columns of $P$ are the vecs of a basis of $F^n$ consisting of eigenvecs of $A$ with eigenvals $\lambda_1, ..., \lambda_n$.
			
		\end{dfn}
		
		
		\begin{thm}[Cayley-Hamilton Theorem]{thm:cayham}{ }
			Let $R$ be a comring and $A \in $Mat$(n; R)$. Then evaluating $\chi_A(x)$ at the matrix $A$ gives zero.
			
		\end{thm}
		
		
		
	%----------------------------_-----------------------------------------------------%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% CHAPTER 5 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%----------------------------------------------------------------------------------%
	
	
	{\footnotesize \textbf{\underline{Inner Product Spaces}}}
		
		\begin{mdfn}[Inner Products And Their Spaces]{mdfn:inprodnspaces}{ }
			Let $V$ be an $\mathbb{R}-$VS. An \textbf{inner product} on $V$ is a mapping $(-, -): V \times V \to \mathbb{R}$ that satisfies, for all $\vec{x}, \vec{y}, \vec{z} \in V$ and $\lambda, \mu \in \mathbb{R}$:
			\begin{enumerate}[label=(\arabic*)]
				\item $(\lambda \vec{x} + \mu \vec{y}, \vec{z}) = \lambda(\vec{x}, \vec{z}) + \mu (\vec{y}, \vec{z})$
				\item $(\vec{x}, \vec{y}) = (\vec{y}, \vec{x})$
				\item $(\vec{x}, \vec{x}) \ge 0$, $=0$ iff $\vec{x} = 0$
			\end{enumerate}
			\begin{itemize}
				\item The \textbf{standard inner product} on $V = \mathbb{R}^n$ is the dot product
				\item A \textbf{real inner product space} is a real vector space endowed with an inner product
				\item $(-, -)$ is a symmetric bilinear form $(1), (2)$ and \textbf{positive definite} $(3)$.
				\item For $V$ a VS over $\mathbb{C}$ the definition is similar, swap $\mathbb{R}$ for $\mathbb{C}$ everywhere; condition $(2)$ reads $(\vec{x}, \vec{y}) = \overline{(\vec{y}, \vec{x})}$; and the standard inner product is $(\vec{v}, \vec{w}) = v_1 \overline{w}_1 + ... + v_n \overline{w}_n$. Note the SIP is not $\mathbb{C}$-linear in the second variable.
				\item In an $\mathbb{R}$ or $\mathbb{C}-$VS, we define the \textbf{length/inner product norm/norm} as $\left\lVert \vec{v} \right\rVert := \sqrt{(\vec{v}, \vec{v})}$
				\item Vectors of length $1$ are \textbf{units}
				\item $\vec{v}, \vec{w}$ are \textbf{orthogonal}, $\vec{v} \perp \vec{w}$, iff $(\vec{v}, \vec{w}) = 0$. Sets $S, T$ are orthogonal, $S \perp T$, if $\vec{v} \perp \vec{w} \forall \vec{v} \in S, \vec{w} \in T$.
				\item A family $(\vec{v}_i)_{i \in I}$ for vecs from an $IP-$space is an \textbf{orthonormal family} if all $\vec{v}_i$ have length $1$ and are pairwise orthogonal (using the Kronecker delta, $(\vec{v}_i, \vec{v}_j) = \delta_{ij}$
				\item An orthonormal family that is a basis is an \textbf{orthonormal basis}. For $V$ an IPS and $(\vec{v}_i)_{i \in I})$ an orthonormal basis, for any $\vec{w} \in V$, $\vec{w} = \sum_{i \in I} (\vec{w}, \vec{v}_i)\vec{v}_i$.
				\item Every finite-dim inner product space has an orthonormal basis
			\end{itemize}
			
		\end{mdfn}
		
		
		\begin{dfn-s}[Orthogonal Complements/Projections]{dfn-s:orthocomplandprojecs}{ }
			Let $V$ be an IPS and $T \subseteq V$. Define $T^\perp = \left\{\vec{v} \in V : \vec{v} \perp \vec{t} \forall \vec{t} \in T\right\}$, the \textbf{orthogonal} to $T$. $T^\perp$ is always a subspace of $V$ even if $T$ is not, and $T^\perp = \left\langle T\right\rangle^\perp.$ \\
			Let $U$ be a finite-dim subspace of an IPS $V$. The space $U^\perp$ is the \textbf{orthogonal complement to $U$}. 
			\begin{itemize}
				\item Every $\vec{v} \in V$ can be written as $\vec{v} = \vec{p} + \vec{r}, \vec{p} \in U$ and $\vec{r} \in U^\perp$.
				\item For $V$ an IPS and $U$ a finite-dim subspace of $V$, $U$ and $U^\perp$ are complementary, ie $V = U \oplus U^\perp$.
			\end{itemize}
			The \textbf{orthogonal projection from $V$ onto $U$} is the mapping $\pi_U : V \to V, \vec{v} = \vec{p} + \vec{r} \mapsto \vec{p}$.
			\begin{itemize}
				\item $\pi_U$ is a linmap with im$(\pi_U) = U$ and $\ker (\pi_U) = U^\perp$
				\item $\pi^2_U = \pi_U$, ie $\pi_U$ is idempotent
				\item If $\{\vec{v}_1, ..., \vec{v}_n\}$ is an orthonormal basis of $U$, then for all $\vec{v} \in V$, $\pi_U(\vec{v}) = \sum_{i=1}^{n}(\vec{v}, \vec{v}_i)\vec{v}_i$.
			\end{itemize}
			
		\end{dfn-s}
		
		
		\begin{mthm}[The Cauchy-Schwarz Inequality]{mthm:caucschw}{ }
			Let $\vec{v}, \vec{w}$ be vectors in an inner product space. Then $\left|(\vec{v}, \vec{w})\right| \le \left\lVert \vec{v}\right\rVert \left\lVert \vec{w}\right\rVert$ with equality iff $\vec{v}, \vec{w}$ are linearly dependent.\\
			For $V$ an IPS, $\vec{v}, \vec{w} \in V$, and $\lambda$ a scalar (ie in the underpinning field):
			\begin{itemize}
				\item $\left\lVert \vec{v}\right\rVert \ge 0$, $=0$ iff $\vec{v} = \vec{0}$.
				\item $\left\lVert \lambda\vec{v}\right\rVert = \left|\lambda\right| \left\lVert \vec{v}\right\rVert$
				\item $\left\lVert \vec{v} + \vec{w}\right\rVert \le \left\lVert \vec{v}\right\rVert + \left\lVert \vec{w}\right\rVert$.
				\item In a real IPS, $\exists$ unique $0 \le \theta \le \pi$ s.t. $\cos(\theta) = \frac{(\vec{v}, \vec{w})}{\left\lVert \vec{v}\right\rVert \left\lVert \vec{w}\right\rVert}$.
			\end{itemize}
			
		\end{mthm}
		
		
		\begin{thm}[Gram-Schmidt Process]{thm:gramschmidt}{ }
			Let $\vec{v}_1, ..., \vec{v}_k$ be linearly ind. vectors in an IPS $V$. Then we can construct an orthonormal family $\vec{w}_1, ..., \vec{w}_k$ such that for all $1 \le i \le k$, $\left\langle \vec{v}_1, ..., \vec{v}_i\right\rangle = \left\langle \vec{w}_1, ..., \vec{w}_i\right\rangle$. If $V$ is finite-dim, we can use the G-S process to transform any basis into an orthonormal basis.\\
			
			\begin{itemize}
				\item For $k = 1$ take $\vec{w}_1 = \frac{\vec{v}_1}{\left\lVert \vec{v}_1\right\rVert}$ (normalise the length)
				\item Then given $\vec{w}_1, ..., \vec{w}_k$; $\vec{w}_{k+1} = \frac{\vec{v}_{k+1} - \sum_{i=1}^{k}(\vec{v}_{k+1}, \vec{w}_i)\vec{w}_i}{\left\lVert\vec{v}_{k+1} - \sum_{i=1}^{k}(\vec{v}_{k+1}, \vec{w}_i)\vec{w}_i\right\rVert}$
			\end{itemize}
			
		\end{thm}
		
		
		\begin{dfn}[Adjoints and Self-Adjointness]{dfn:adjnt}{ }
			Let $V$ be an IPS. Two endomorphisms $T, S : V \to V$ are \textbf{adjoint} if for all $\vec{v}, \vec{w} \in V$, $(T \vec{v}, \vec{w}) = (\vec{v}, S\vec{w})$. In this case $S = T*$ and $S$ is the \textbf{adjoint of $T$}.
			\begin{itemize}
				\item Let $V$ be a finite-dim IPS and $T: V \to V$ an endo. Then $T*$ exists an is unique.
				\item An endo. of an IPS $T : V \to V$ is \textbf{self-adjoint} if $T = T*$.
				\item Let $T : V \to V$ be an self-adjoint linmap on an IPS $V$. Then:
					\begin{enumerate}[label=(\arabic*)]
						\item Every eigenvalue of $T$ is real
						\item If $\lambda, \mu$ are distinct eigvals of $T$ with eigenvecs $\vec{v}, \vec{w}$, then $(\vec{v}, \vec{w}) = 0$
						\item T has an eigenvalue.
					\end{enumerate}
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{mthm}[The Spectral Theorem]{mthm:}{ }
			Let $V$ be a finite-dim IPS and $T : V \to V$ be a self-adjoint linmap. Then $V$ has an orthonormal basis consisting of eigenvectors of $T$.
			\begin{itemize}
				\item Let $A \in $Mat$(n, \mathbb{R})$ be symmetric. Then $\exists$ orthogonal $P \in $Mat$(n, \mathbb{R})$ such that $P^TAP = P^{-1}AP = $diag$(\lambda_1, ..., \lambda_n)$, where the $\lambda_i$ are the (necessarily real) eigvals of $A$ repeated according to their multiplicity as roots of $\chi_A$.
				\item Let $A \in $Mat$(n, \mathbb{C})$ be hermitian. Then $\exists$ unitary $P \in $Mat$(n, \mathbb{C})$ s.t. $\overline{P}^T A P = P^{-1} AP = $diag$(\lambda_1, ..., \lambda_n)$, where the $\lambda_i$ are the (necessarily real) eigvals of $A$ repeated according to their multiplicity as roots of $\chi_A$.
			\end{itemize}
			
		\end{mthm}
		
		
		
	%----------------------------------------------------------------------------------%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% CHAPTER 6 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%----------------------------------------------------------------------------------%
	
	
	{\footnotesize \textbf{\underline{Jordan Normal Form}}}
		
		{\tiny In this section assume that $F$ is an algebraically closed field}
		
		
		\begin{dfn}[Jordan Blocks]{dfn:jordblocks}{ }
			Let $r \in \mathbb{N}$. Then 
			\begin{itemize}
				\item The \textbf{nilpotent Jordan block of size $r$} is the $(r \times r)$ matrix $J(r)$ defined by the rule $J(r)_{ij} = 1$ for $j = i + i$ and $=0$ otherwise
				\item The \textbf{Jordan block of size $r$ and eigenvalue $\lambda$}, for $\lambda \in F$, is the $(r \times r)$ matrix $J(r, \lambda)$ defined by the rule $J(r, \lambda) = \lambda I_r + J(r) = D + N$ (with $DN = ND$).
			\end{itemize}
			
		\end{dfn}
		
		
		\begin{mthm}[Jordan Normal Form]{mthm:jordnormform}{ }
			Let $F$ be an algebraically closed field, $V$ be a finite-dim VS, and $\phi : V \to V$ an endomorphism with characteristic polynomial $\chi_\phi(x) = (x - \lambda_1)^{a_1}(x-\lambda_2)^{a_2}...(x-\lambda_s)^{a_s} \in F[x]$, $a_i \ge 1, \sum_{i=1}^{s} = n$ for distinct $\lambda_1, ..., \lambda_s \in F$.\\
			Then there exists an ordered basis $\mathcal{B}$ of $V$ such that the matrix of $\phi$ with respect to the basis $\mathcal{B}$ is block diagonal with Jordan blocks on the diagonal. That is, $_\mathcal{B}[\phi]_\mathcal{B} = $diag$(J(r_{11}, \lambda_1), ..., J(r_{1m_1}, \lambda_1), J(r_{21}, \lambda_2), ..., J(r_{sm_s}, \lambda_s))$ with $r_{11}, ..., r_{1m_1}, r_{21}, ..., r_{sm_s} \ge 1$ such that $a_i = r_{i1} + r_{i2} + ... + r_{im_i}, (1 \le i \le s)$.
			
		\end{mthm}
		
		
		\begin{rem}[The Proof of Jordan Normal Form]{rem:jordnormformproof}
		
			%TO-DO - HONOURS ALGEBRA 1 - 0 LOUIS
			
		\end{rem}
		
		
		
		
		
		
		
	\end{multicols*}
	
	
\end{document}
